---
title: "Dunlop growth rates - data clean up and summary statistics"
author: "Cassandra Wattenburger"
date: "11/10/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(results = "show")
```

**Goal of this script:** Explore the properties of the data and descriptive statistics of in-situ growth rates.

# Import libraries

```{r}
library(tidyverse)
library(cowplot)
library(phyloseq)
library(lmerTest)
library(emmeans)

sessionInfo()

rm(list=ls())
```

# Load data and clean up

There are inconsistencies in the data structure, variable names, and metadata that need to be cleaned up for easier manipulation and analysis.

Internal-standard normalized microcosm and field community data:

```{r}
# Original, rarefied phyloseq object (not normalized)
load("rdata.files/gr_physeq.rarefy.RData")
tax = as.data.frame(tax_table(physeq.rarefy)) %>% rownames_to_column(var="ASV") # taxonomy extract

# Microcosms
load("rdata.files/gr_microcosm.norm.tax.RData")
ucosms = microcosm.norm.tax

# Replicates
ucosm.rep1 = seq(81, 260, by=3)
ucosm.rep2 = seq(82, 260, by=3)
ucosm.rep3 = seq(83, 260, by=3)

# Clean up
ucosms = ucosms %>% 
  mutate(Sample = str_remove(SampleID, "sa"), Sample = parse_number(Sample)) %>% # convert sample numbers into numeric (easier manipulation)
  mutate(Replicate = ifelse(Sample %in% ucosm.rep1, 1, ifelse(Sample %in% ucosm.rep2, 2, ifelse(Sample %in% ucosm.rep3, 3, NA)))) %>% # add replicates
  select(Sample, Amendment=Cammend, Soil, Day=Days, Replicate, norm_abund=value, ASV=variable, Domain, Phylum, Class, Order, Family, Genus) # rename/reorder

# Field
load("rdata.files/gr_field.norm.RData")
field = field.norm

# Replicates
field.rep1 = seq(1, 80, by=4)
field.rep2 = seq(2, 80, by=4)
field.rep3 = seq(3, 80, by=4)
field.rep4 = seq(4, 80, by=4)

# Clean up
field = field %>%
  mutate(Sample = str_remove(SampleID, "sa"), Sample = parse_number(Sample)) %>% # convert sample numbers into numeric (easier manipulation)
  pivot_longer(9:33122, names_to = "ASV", values_to="norm_abund") %>% # tidy format
  filter(Soil %in% c("C3", "S17")) %>% # only these plots are relevant
  mutate(Replicate = ifelse(Sample %in% field.rep1, 1, ifelse(Sample %in% field.rep2, 2, ifelse(Sample %in% field.rep3, 3, ifelse(Sample %in% field.rep4, 4, NA))))) %>% # add replciates
  select(Sample, Soil, Block, Replicate, ASV, norm_abund) %>% # rename/reorder
  left_join(tax, by="ASV") # add taxonomy
```

Number of ASVs/phyla in dataset:

```{r}
# Extract from phyloseq
count = as.data.frame(otu_table(physeq.rarefy))
meta = data.frame(sam_data(physeq.rarefy))

# Remove unused samples
## Only microcosm adn S17/C3 from field relevant
meta.rm = meta %>% subset(Soil == "C3" | Soil == "S17")

count.rm = count %>%
  t() %>%
  as.data.frame() %>%
  merge(meta.rm, by=0) %>%
  column_to_rownames(var="Row.names") %>%
  select(everything(), -(SampleID:Experiment))

# Remove absent ASVs
count.rm0 = colSums(count.rm, dims=1)
count.rm0 = count.rm0[count.rm0 > 0]

# Total ASVs
length(count.rm0)
asvs = names(count.rm0)

# Total phyla
tax.rm0 = tax[tax$ASV %in% asvs,]
phyla = unique(tax.rm0$Phylum)
```

Estimated in-situ growth rates:

```{r}
gr.est = readRDS("rdata.files/gr_gr.final_rm0.rds")

# Clean up
gr.est = gr.est %>% 
  left_join(tax, by="ASV") %>% # add taxonomy
  select(Soil, Amendment, Replicate, label, Slope, k, g, Start, End, Length, ASV, Domain, Phylum, Class, Order, Family, Genus) # rename/reorder
```

Fix start/end values in growth rate data:

Currently these represent the rows of the dataframe where the growth rate estimate started and ended, NOT the day of incubation.

```{r}
# Remove 0s from normalized ucosms data
## Ensures that row number aligns with correct day information, because we removed 0s for growth rate estimation (see script 03)
ucosm.rm0 = filter(ucosms, norm_abund > 0)

# Gather start and end of growth data for each ASV in each treatment
gr.est.new = data.frame()
for (s in unique(gr.est$Soil)) {
  for (a in unique(gr.est[gr.est$Soil==s,]$Amendment)) {
    for (r in unique(gr.est[gr.est$Soil==s & gr.est$Amendment==a,]$Replicate)) {
      for (i in unique(gr.est[gr.est$Soil==s & gr.est$Amendment==a & gr.est$Replicate==r,]$ASV)) {
        gr.sub = filter(gr.est, Soil==s & Amendment==a & Replicate==r & ASV==i)
        start.tp = gr.sub$Start # time point NOT day
        end.tp = gr.sub$End # time point NOT day
        ucosm.sub = filter(ucosm.rm0, Soil==s & Amendment==a & Replicate==r & ASV==i) %>%
          arrange(Day) # need to order by day so start and end select correct day value
        start.day = ucosm.sub[start.tp,]$Day # actual start day
        end.day = ucosm.sub[end.tp,]$Day # actual end day
        
        # Rebuild gr.est dataframe
        newrow = data.frame("Soil"=s, "Amendment"=a, "Replicate"=r,
                            "ASV"=i, "Domain"=gr.sub$Domain, "Phylum"=gr.sub$Phylum,
                            "Class"=gr.sub$Class, "Order"=gr.sub$Order, "Family"=gr.sub$Family,
                            "Genus"=gr.sub$Genus,
                            "k"=gr.sub$k, "g"=gr.sub$g, "start_tp"=start.tp, "end_tp"=end.tp, "start_day"=start.day, "end_day"=end.day)
        gr.est.new = est.new = rbind(gr.est.new, newrow)
      }
    }
  }
}

```


```{r, eval=FALSE}
# Save cleaned datasets for use in future scripts
saveRDS(ucosms, file="rdata.files/ucosm.norm.cleaned.rds")
saveRDS(field, file="rdata.files/field.norm.cleaned.rds")
saveRDS(gr.est.new, file="rdata.files/gr.final.cleaned.rds")
```


# Overview of communities

Number of ASVs/phyla

```{r}
# Microcosms
ucosms %>% filter(norm_abund != 0) %>% distinct(ASV) %>% nrow() # total ASVs (not including 0 abundance)

ucosms %>%  # average ASVs per sample
  filter(norm_abund != 0) %>% 
  group_by(Sample) %>% 
  summarize(num_ASV = length(unique(ASV))) %>% 
  colMeans()

ucosms %>% # total phyla, not counting 0 abundance ASVs, or uncertain taxonomic placement
  filter(norm_abund != 0) %>% 
  mutate(Phylum = str_remove(Phylum, "putative "), Phylum = str_remove(Phylum, "unclassified ")) %>%
  distinct(Phylum) %>% 
  nrow()

# Field
field %>% filter(norm_abund != 0) %>% distinct(ASV) %>% nrow() # total ASVs (not including 0 abundance)

field %>%  # average ASVs per sample
  filter(norm_abund != 0) %>% 
  group_by(Sample) %>% 
  summarize(num_ASV = length(unique(ASV))) %>% 
  colMeans()

field %>% # total phyla, not counting 0 abundance ASVs, or uncertain taxonomic placement
  filter(norm_abund != 0) %>% 
  mutate(Phylum = str_remove(Phylum, "putative "), Phylum = str_remove(Phylum, "unclassified ")) %>% 
  distinct(Phylum) %>% 
  nrow() 

# Microcosms and field together (full dataset)
ucosms %>% bind_rows(field) %>% filter(norm_abund != 0) %>% distinct(ASV) %>% nrow() # total ASVs (not including 0 abundance)

ucosms %>% # average number of ASVs
  bind_rows(field) %>% 
  filter(norm_abund != 0) %>% 
  group_by(Sample) %>% 
  summarize(num_ASV = length(unique(ASV))) %>% 
  colMeans()

ucosms %>% # total phyla, not counting 0 abundance ASVs, or uncertain taxonomic placement
  bind_rows(field) %>% 
  filter(norm_abund != 0) %>% 
  mutate(Phylum = str_remove(Phylum, "putative "), Phylum = str_remove(Phylum, "unclassified ")) %>%
  distinct(Phylum) %>% 
  nrow() 

# Estimated growth rates
gr.est %>% distinct(ASV) %>% nrow() # total ASVs (not including 0 abundance)

gr.est %>% # total phyla, not counting 0 abundance ASVs, not counting putative/unclassified
  mutate(Phylum = str_remove(Phylum, "putative "), Phylum = str_remove(Phylum, "unclassified ")) %>% 
  distinct(Phylum) %>% 
  nrow()

# Number of estimates per phylum
gr.est %>% 
  mutate(Phylum = str_remove(Phylum, "putative ")) %>% # roll putative into the phylum
  distinct(ASV, Phylum) %>%
  group_by(Phylum) %>%
  summarize(total = n()) %>%
  arrange(-total)

# Number of estimates per treatment
gr.est %>%
  group_by(Soil, Amendment) %>%
  distinct(ASV) %>%
  summarize(total = n())
```

Most abundant phyla

```{r}
# Microcosms (tp0)
ucosms %>% 
  filter(Day==0) %>%
  group_by(Replicate, Phylum) %>% 
  summarize(total_abund = sum(norm_abund)) %>%
  ggplot(aes(y=fct_reorder(Phylum, -total_abund), x=total_abund)) +
  geom_boxplot() +
  labs(title="Microcosms at TP0", x="Total normalized abundance", y="Phylum") +
  theme_test()

# Field
field %>% 
  group_by(Replicate, Phylum) %>% 
  summarize(total_abund = sum(norm_abund)) %>%
  ggplot(aes(y=fct_reorder(Phylum, -total_abund), x=total_abund)) +
  geom_boxplot() +
  labs(title="Field", x="Total normalized abundance", y="Phylum") +
  theme_test()

# Growth rate estimates
gr.est %>% 
  group_by(Replicate, Phylum) %>%
  summarize(num_ASV = length(ASV)) %>%
  ggplot(aes(y=fct_reorder(Phylum, -num_ASV), x=num_ASV)) +
  geom_boxplot() +
  labs(title="Growth estimates", x="Total ASVs", y="Phylum") +
  theme_test()

# Count total ASVs in overall data
gr.est %>% group_by(Phylum) %>% count(Phylum) %>% arrange(-n)
```

Top 5 most abundant phyla in all data categories are Proteobacteria, Actinobacteria, Acidobacteria, Verrucomicrobia, and Planctomycetes in that order.


