---
title: "Dunlop growth rates - data exploration and summary statistics"
author: "Cassandra Wattenburger"
date: "11/10/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(results = "show")
```

**Goal of this script:** Explore the properties of the data and descriptive statistics of in-situ growth rates.

# Import libraries

```{r}
library(tidyverse)
library(phyloseq)

sessionInfo()

rm(list=ls())
```

# Load data and clean up

There are inconsistencies in the data structure, variable names, and metadata that need to be cleaned up for easier manipulation and analysis.

Internal-standard normalized microcosm and field community data:

```{r}
# Original, rarefied phyloseq object (not normalized)
load("rdata.files/gr_physeq.rarefy.RData")
tax = as.data.frame(tax_table(physeq.rarefy)) %>% rownames_to_column(var="ASV") # taxonomy extract

# Microcosms
load("rdata.files/gr_microcosm.norm.tax.RData")
ucosms = microcosm.norm.tax

# Replicates
ucosm.rep1 = seq(81, 260, by=3)
ucosm.rep2 = seq(82, 260, by=3)
ucosm.rep3 = seq(83, 260, by=3)

# Clean up
ucosms = ucosms %>% 
  mutate(Sample = str_remove(SampleID, "sa"), Sample = parse_number(Sample)) %>% # convert sample numbers into numeric (easier manipulation)
  mutate(Replicate = ifelse(Sample %in% ucosm.rep1, 1, ifelse(Sample %in% ucosm.rep2, 2, ifelse(Sample %in% ucosm.rep3, 3, NA)))) %>% # add replicates
  select(Sample, Amendment=Cammend, Soil, Day=Days, Replicate, norm_abund=value, ASV=variable, Domain, Phylum, Class, Order, Family, Genus) # rename/reorder

# Field
load("rdata.files/gr_field.norm.RData")
field = field.norm

# Replicates
field.rep1 = seq(1, 80, by=4)
field.rep2 = seq(2, 80, by=4)
field.rep3 = seq(3, 80, by=4)
field.rep4 = seq(4, 80, by=4)

# Clean up
field = field %>%
  mutate(Sample = str_remove(SampleID, "sa"), Sample = parse_number(Sample)) %>% # convert sample numbers into numeric (easier manipulation)
  pivot_longer(9:33122, names_to = "ASV", values_to="norm_abund") %>% # tidy format
  filter(Soil %in% c("C3", "S17")) %>% # only these plots are relevant
  mutate(Replicate = ifelse(Sample %in% field.rep1, 1, ifelse(Sample %in% field.rep2, 2, ifelse(Sample %in% field.rep3, 3, ifelse(Sample %in% field.rep4, 4, NA))))) %>% # add replciates
  select(Sample, Soil, Block, Replicate, ASV, norm_abund) %>% # rename/reorder
  left_join(tax, by="ASV") # add taxonomy
```

Estimated in-situ growth rates:

```{r}
gr.est = readRDS("rdata.files/gr_gr.final_rm0.rds")

# Clean up
gr.est = gr.est %>% 
  left_join(tax, by="ASV") %>% # add taxonomy
  select(Soil, Amendment, Replicate, label, Slope, k, g, Start, End, Length, ASV, Domain, Phylum, Class, Order, Family, Genus) # rename/reorder
```

```{r, eval=FALSE}
# Save cleaned datasets for use in future scripts
saveRDS(ucosms, file="rdata.files/ucosm.norm.cleaned.rds")
saveRDS(field, file="rdata.files/field.norm.cleaned.rds")
saveRDS(gr.est, file="rdata.files/gr.final.cleaned.rds")
```

# Overview of community diversity

Number of ASVs/phyla

```{r}
# Microcosms
ucosms %>% filter(norm_abund != 0) %>% distinct(ASV) %>% nrow() # total ASVs (not including 0 abundance)

ucosms %>%  # average ASVs per sample
  filter(norm_abund != 0) %>% 
  group_by(Sample) %>% 
  summarize(num_ASV = length(unique(ASV))) %>% 
  colMeans()

ucosms %>% # total phyla, not counting 0 abundance ASVs, or uncertain taxonomic placement
  filter(norm_abund != 0) %>% 
  mutate(Phylum = str_remove(Phylum, "putative "), Phylum = str_remove(Phylum, "unclassified ")) %>%
  distinct(Phylum) %>% 
  nrow()

# Field
field %>% filter(norm_abund != 0) %>% distinct(ASV) %>% nrow() # total ASVs (not including 0 abundance)

field %>%  # average ASVs per sample
  filter(norm_abund != 0) %>% 
  group_by(Sample) %>% 
  summarize(num_ASV = length(unique(ASV))) %>% 
  colMeans()

field %>% # total phyla, not counting 0 abundance ASVs, or uncertain taxonomic placement
  filter(norm_abund != 0) %>% 
  mutate(Phylum = str_remove(Phylum, "putative "), Phylum = str_remove(Phylum, "unclassified ")) %>% 
  distinct(Phylum) %>% 
  nrow() 

# Microcosms and field together (full dataset)
ucosms %>% bind_rows(field) %>% filter(norm_abund != 0) %>% distinct(ASV) %>% nrow() # total ASVs (not including 0 abundance)

ucosms %>% # average number of ASVs
  bind_rows(field) %>% 
  filter(norm_abund != 0) %>% 
  group_by(Sample) %>% 
  summarize(num_ASV = length(unique(ASV))) %>% 
  colMeans()

ucosms %>% # total phyla, not counting 0 abundance ASVs, or uncertain taxonomic placement
  bind_rows(field) %>% 
  filter(norm_abund != 0) %>% 
  mutate(Phylum = str_remove(Phylum, "putative "), Phylum = str_remove(Phylum, "unclassified ")) %>%
  distinct(Phylum) %>% 
  nrow() 

# Estimated growth rates
gr.est %>% distinct(ASV) %>% nrow() # total ASVs (not including 0 abundance)

gr.est %>% # total phyla, not counting 0 abundance ASVs, or uncertain taxonomic placement
  mutate(Phylum = str_remove(Phylum, "putative "), Phylum = str_remove(Phylum, "unclassified ")) %>% 
  distinct(Phylum) %>% 
  nrow() 
```

Most abundant phyla

```{r}
# Microcosms (tp0)
ucosms %>% 
  filter(Day==0) %>%
  group_by(Replicate, Phylum) %>% 
  summarize(total_abund = sum(norm_abund)) %>%
  ggplot(aes(y=fct_reorder(Phylum, -total_abund), x=total_abund)) +
  geom_boxplot() +
  labs(title="Microcosms at TP0", x="Total normalized abundance", y="Phylum") +
  theme_test()

# Field
field %>% 
  group_by(Replicate, Phylum) %>% 
  summarize(total_abund = sum(norm_abund)) %>%
  ggplot(aes(y=fct_reorder(Phylum, -total_abund), x=total_abund)) +
  geom_boxplot() +
  labs(title="Field", x="Total normalized abundance", y="Phylum") +
  theme_test()

# Growth rate estimates
gr.est %>% 
  group_by(Replicate, Phylum) %>%
  summarize(num_ASV = length(ASV)) %>%
  ggplot(aes(y=fct_reorder(Phylum, -num_ASV), x=num_ASV)) +
  geom_boxplot() +
  labs(title="Growth estimates", x="Total ASVs", y="Phylum") +
  theme_test()

# Count total ASVs in overall data
gr.est %>% group_by(Phylum) %>% count(Phylum) %>% arrange(-n)
```

Top 5 most abundant phyla in all data categories are Proteobacteria, Actinobacteria, Acidobacteria, Verrucomicrobia, and Planctomycetes in that order.

# Growth estimate summaries

Start of growth

```{r}
gr.est %>%
  group_by(Soil, Amendment, Replicate) %>%
  summarize(avg_start = mean(Start)) %>%
  ggplot(aes(x=Soil, y=avg_start, color=Amendment)) +
  geom_boxplot() +
  labs(title="Start of growth window", y="Day") +
  theme_test()
```

End of growth window

```{r}
gr.est %>%
  group_by(Soil, Amendment, Replicate) %>%
  summarize(avg_end = mean(End)) %>%
  ggplot(aes(x=Soil, y=avg_end, color=Amendment)) +
  geom_boxplot() +
  labs(title="End of growth window", y="Day") +
  theme_test()
```

Length of growth window

```{r}
gr.est %>%
  group_by(Soil, Amendment, Replicate) %>%
  summarize(avg_len = mean(Length)) %>%
  ggplot(aes(x=Soil, y=avg_len, color=Amendment)) +
  geom_boxplot() +
  labs(title="Length of growth window", y="Day") +
  theme_test()
```


Specific growth rates

```{r}
gr.est %>%
  group_by(Soil, Amendment, Replicate) %>%
  summarize(avg_k = mean(k)) %>%
  ggplot(aes(x=Soil, y=avg_k, color=Amendment)) +
  geom_boxplot() +
  labs(title="Specific growth rate", y="Specific growth rate (per Day)") +
  theme_test()
```

Create table:

```{r}
gr.est %>%
  group_by(Soil, Amendment) %>%
  summarize(avg_k = mean(k), sd_k = sd(k), avg_start = mean(Start), sd_start = sd(Start), avg_end = mean(End), sd_end = sd(End), avg_len = mean(Length), sd_len = sd(Length))
```

Generation time (note that short generation times are fast growth, long generation times are slow growth)

```{r}
gr.est %>% summarize(min_g=min(g), max_g=max(g), mean_g=mean(g))
```


# Overlapping estimates between treatments

```{r}
# Isolate treatment ASV names
gr.C3.asv = gr.est %>% filter(Soil=="C3") %>% distinct(ASV)
gr.S17.asv = gr.est %>% filter(Soil=="S17") %>% distinct(ASV)
gr.C3y.asv = gr.est %>% filter(Soil=="C3", Amendment=="Y") %>% distinct(ASV)
gr.C3n.asv = gr.est %>% filter(Soil=="C3", Amendment=="N") %>% distinct(ASV)
gr.S17y.asv = gr.est %>% filter(Soil=="S17", Amendment=="Y") %>% distinct(ASV)
gr.S17n.asv = gr.est %>% filter(Soil=="S17", Amendment=="N") %>% distinct(ASV)

# Overlap
inner_join(gr.C3.asv, gr.S17.asv) %>% nrow()
inner_join(gr.C3y.asv, gr.C3n.asv) %>% nrow()
inner_join(gr.S17y.asv, gr.S17n.asv) %>% nrow()
```


# Biogeochemical data analysis

Analyze SOM and pH data collected from in-field samples.

### SOM

```{r}
# Import data
som = read_csv("biogeochem_data/growthrate_insitu_som.csv")
som = som %>% 
  filter((Soil=="S" & Year==17) | (Soil=="C" & Year==3)) %>%
  mutate(som = (Preignition - Postignition) / (Preignition - Crucible) * 100, # calculate % SOM
         Soil = parse_factor(Soil),
         Soil = recode_factor(Soil, C="Cropped", S="Successional")) # clean up

som %>% 
  ggplot(aes(x=Soil, y=som)) +
  geom_boxplot() +
  labs(y="Soil organic matter (%)") +
  theme_test()
```

**Statistics**

```{r}
# Trying two models, one with interactive effect with block
som.lm1 = lm(som ~ Soil + Block, data=som)
som.lm2 = lm(som ~ Soil*Block, data=som)

# Compare models
summary(som.lm1)
summary(som.lm2)
anova(som.lm1, som.lm2) # comparison
```

The model that accounts for an interaction with block fits best.

```{r}
hist(resid(som.lm2)) # assess nomrality assumption
plot(predict(som.lm2), resid(som.lm2)) # assess equal variances assumption
summary(som.lm2)
```

There's a significant difference between soil som that depends on the block.

Plot block effect:

```{r}
som %>%
  ggplot(aes(x=Soil, y=som)) +
  geom_boxplot() +
  facet_wrap(~Block) +
  labs(x="", y="Soil organic matter (%)", title="Block effect") +
  theme_test()
```

WOW! Block 2 is downslope of block 1, so runoff may be driving SOM accumulation in block 2.


### pH

```{r}
# Import data
ph = read_csv("biogeochem_data/growthrate_insitu_ph.csv")
ph = ph %>% 
  filter((Soil=="S" & Year==17) | (Soil=="C" & Year==3)) %>%
  mutate(Soil = parse_factor(Soil),
         Soil = recode_factor(Soil, C="Cropped", S="Successional")) # clean up

ph %>% 
  ggplot(aes(x=Soil, y=pH)) +
  geom_boxplot() +
  labs(x="") +
  theme_test()
```


**Statistics**

```{r}
# Trying two models, one with interactive effect with block
ph.lm1 = lm(pH ~ Soil + Block, data=ph)
ph.lm2 = lm(pH ~ Soil*Block, data=ph)

# Compare
summary(ph.lm1)
summary(ph.lm2)
anova(ph.lm1, ph.lm2)
```

Model that allows interaction with block explains much more variance (see adjusted R-squared). ANOVA comparison also finds significant difference between the models.

```{r}
hist(resid(ph.lm2)) # assess nomrality assumption
plot(predict(ph.lm2), resid(ph.lm2)) # assess equal variances assumption
summary(ph.lm2)
```

There is a significant difference in the pH of soils that depends on the block. Expected based on results from SOM data.

Plot block effect:

```{r}
ph %>%
  ggplot(aes(x=Soil, y=pH)) +
  geom_boxplot() +
  facet_wrap(~Block) +
  labs(x="", y="Soil organic matter (%)", title="Block effect") +
  theme_test()
```

Block 2 seems to have lower pH overall compared to block 1. 

### Make table of results

```{r}
# Merge SOM and pH data
som.ph = full_join(som, ph) %>%
  group_by(Soil) %>%
  summarize(som.avg = mean(som, na.rm=TRUE), som.sd = sd(som, na.rm=TRUE),
            ph.avg = mean(pH, na.rm=TRUE), ph.sd = sd(pH, na.rm=TRUE))
som.ph
```


