---
title: "Dunlop growth rates - habitat"
author: "Cassandra Wattenburger"
date: "11/10/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(results = "show")
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

**Goal of this script:** Investigate the effects of soil habitat on in-situ growth rates

# Import libraries

```{r}
library(tidyverse)
library(cowplot)
library(data.table)
library(phyloseq)
library(lmerTest)

sessionInfo()

rm(list=ls())
```


# Biogeochemical data analyses

Analyze SOM and pH data collected from in-field samples.


### SOM

```{r}
# Import data
som = read_csv("biogeochem_data/growthrate_insitu_som.csv")
som = som %>% 
  filter((Soil=="S" & Year==17) | (Soil=="C" & Year==3)) %>%
  mutate(som = (Preignition - Postignition) / (Preignition - Crucible) * 100, # calculate % SOM
         Soil = parse_factor(Soil),
         Soil = recode_factor(Soil, C="Cropped", S="Successional")) # clean up

som %>% 
  ggplot(aes(x=Soil, y=som)) +
  geom_boxplot() +
  labs(y="Soil organic matter (%)") +
  theme_test()

som %>% 
  ggplot(aes(x=Soil, y=som)) +
  geom_boxplot() +
  facet_wrap(~Block) +
  labs(y="Soil organic matter (%)") +
  theme_test()
```

Strong block effect.

**Statistics**

Linear regression
* Soil and Block as fixed effects with interaction term

```{r}
som.lm = lm(som ~ Soil* Block, data=som)

hist(resid(som.lm))
plot(resid(som.lm), predict(som.lm))

som.lm2 = lm(log(som) ~ Soil* Block, data=som)

hist(resid(som.lm2))
plot(resid(som.lm2), predict(som.lm2))
```

Neither of the residuals look like they fit the assumptions of the linear model... I'll do a welch's t-tests instead, since it looks like unequal variances seem to be the issue. I can't add interaction term, so we'll do a separate test for each block to account for the interaction.

Welch's t-tests

```{r}
t.test(som ~ Soil, data=som[som$Block==1,], var.equal=FALSE) # block 1 only
t.test(som ~ Soil, data=som[som$Block==2,], var.equal=FALSE) # block 2 only
```

SOM is signficant between soils in block 2 but not block 1. Likely due to block 2 being downslope of block 1.


### pH

```{r}
# Import data
ph = read_csv("biogeochem_data/growthrate_insitu_ph.csv")
ph = ph %>% 
  filter((Soil=="S" & Year==17) | (Soil=="C" & Year==3)) %>%
  mutate(Soil = parse_factor(Soil),
         Soil = recode_factor(Soil, C="Cropped", S="Successional")) # clean up

ph %>% 
  ggplot(aes(x=Soil, y=pH)) +
  geom_boxplot() +
  labs(x="") +
  theme_test()

ph %>% 
  ggplot(aes(x=Soil, y=pH)) +
  geom_boxplot() +
  facet_wrap(~Block) +
  labs(x="") +
  theme_test()
```

Also looks like there is a strong block effect here.


**Statistics**

Linear regression
* Soil and Block as fixed effects with interaction term

```{r}
ph.lm = lm(pH ~ Soil*Block, data=ph)

hist(resid(ph.lm)) # check normality
plot(resid(ph.lm), predict(ph.lm)) # check variances

summary(ph.lm)
```

There is a significant soil, block, and interaction between block and soil for pH.


### Make table of results

```{r}
# Merge SOM and pH data
som.ph = full_join(som, ph) %>%
  group_by(Soil) %>%
  summarize(som.avg = mean(som, na.rm=TRUE), som.sd = sd(som, na.rm=TRUE),
            ph.avg = mean(pH, na.rm=TRUE), ph.sd = sd(pH, na.rm=TRUE))
som.ph
```

Import estimated growth rates data:

```{r}
gr.est = readRDS("rdata.files/gr.final.cleaned.rds")
```

# Growth properties

```{r}
# Calculate length of growth based on start and end
gr.est = mutate(gr.est, length_days = end_day - start_day)

# Average across ASVs for each replicate
gr.est.avg = gr.est %>%
  group_by(Soil, Amendment, Replicate) %>%
  summarize(avg_start = mean(start_day),
            avg_end = mean(end_day),
            avg_length = mean(length_days),
            avg_k = mean(k))
```


### Start of growth

```{r}
gr.est.avg %>%
  ggplot(aes(x=Soil, y=avg_start, color=Amendment)) +
  geom_boxplot() +
  labs(title="Start of growth", y="Day") +
  theme_test()
```

**Statistics**

Try ANOVA.

```{r}
strt.aov = aov(avg_start ~ Soil + Amendment + Soil:Amendment, data=gr.est.avg)
hist(resid(strt.aov)) # check normality
plot(predict(strt.aov), resid(strt.aov)) # check residuals
```

Everything looks good.

```{r}
summary(strt.aov)
```

Start of growth is affected by soil and amendment.

Post hoc: 

Only interested in certain contrasts, ie doesn't make sense to compare cropped water-amended to successional C-amended.

Using Student's t-tests with Benjamini-Hochberg correction.

```{r}
# Contrasts
start.ttest1 = t.test(log(avg_start) ~ Amendment, data=gr.est.avg[gr.est.avg$Soil=="C3",], var.equal = TRUE)
start.ttest2 = t.test(log(avg_start) ~ Amendment, data=gr.est.avg[gr.est.avg$Soil=="S17",], var.equal = TRUE)
start.ttest3 = t.test(log(avg_start) ~ Soil, data=gr.est.avg[gr.est.avg$Amendment=="Y",], var.equal = TRUE)
start.ttest4 = t.test(log(avg_start) ~ Soil, data=gr.est.avg[gr.est.avg$Amendment=="N",], var.equal = TRUE)

# Adjust p-values
start.pvals = c(start.ttest1$p.value, start.ttest2$p.value, start.ttest3$p.value, start.ttest4$p.value)
start.adjpvals = p.adjust(start.pvals, method="BH", n=length(start.pvals))

# Results
start.results = data.frame(contrast = c("Cropped, C vs W", "Successional, C vs W", "C-amended, cropped vs succ.", "Water, cropped vs succ."), pvals = start.adjpvals)
start.results
```

No significance.


### End of growth window

```{r}
gr.est.avg %>%
  ggplot(aes(x=Soil, y=avg_end, color=Amendment)) +
  geom_boxplot() +
  labs(title="End of growth", y="Day") +
  theme_test()
```

**Statistics**

```{r}
end.aov = aov(avg_end ~ Soil + Amendment + Soil:Amendment, data=gr.est.avg)
hist(resid(end.aov)) # check normality
```

Try log transform

```{r}
end.aov2 = aov(log(avg_end) ~ Soil + Amendment + Soil:Amendment, data=gr.est.avg)
hist(resid(end.aov2)) # check normality
```

Not better, stick with original data.

```{r}
plot(predict(end.aov), resid(end.aov))
summary(end.aov)
```

End of growth is not affected by soil or amendment.


### Length of growth window

```{r}
gr.est.avg %>%
  ggplot(aes(x=Soil, y=avg_length, color=Amendment)) +
  geom_boxplot() +
  labs(title="Length of growth", y="Day") +
  theme_test()
```

**Statistics**

```{r}
length.aov = aov(avg_length ~ Soil + Amendment + Soil*Amendment, data=gr.est.avg)
hist(resid(length.aov)) # check normality
plot(predict(length.aov), resid(length.aov)) # check residuals
summary(length.aov)
```

Length of growth period not significant.


### Specific growth rates

```{r}
gr.est.avg %>%
  ggplot(aes(x=Soil, y=log(avg_k), color=Amendment)) +
  geom_boxplot() +
  labs(title="Specific growth rate", y="Specific growth rate, ln (per Day)") +
  theme_test()
```


**Statistics**

```{r}
k.aov = lm(avg_k ~ Soil + Amendment + Soil:Amendment, data=gr.est.avg)
hist(resid(k.aov))
```

Try log transform.

```{r}
k.aov2 = aov(log(avg_k) ~ Soil + Amendment + Soil:Amendment, data=gr.est.avg)
hist(resid(k.aov2))
plot(predict(k.aov2), resid(k.aov2))
summary(k.aov2)
```

No significance found.

Generation time summary:

```{r}
gr.est %>% summarize(min_g=min(g), max_g=max(g), mean_g=mean(g))
```

### Create table of results:

```{r}
gr.est %>%
  group_by(Soil, Amendment) %>%
  summarize(avg_k = mean(k), sd_k = sd(k), avg_start = mean(start_day), sd_start = sd(start_day), avg_end = mean(end_day), sd_end = sd(end_day), avg_len = mean(length_days), sd_len = sd(length_days))
```


# Overlapping estimates between treatments

```{r}
# Isolate treatment ASV names
gr.C3.asv = gr.est %>% filter(Soil=="C3") %>% distinct(ASV)
gr.S17.asv = gr.est %>% filter(Soil=="S17") %>% distinct(ASV)
gr.C3y.asv = gr.est %>% filter(Soil=="C3", Amendment=="Y") %>% distinct(ASV)
gr.C3n.asv = gr.est %>% filter(Soil=="C3", Amendment=="N") %>% distinct(ASV)
gr.S17y.asv = gr.est %>% filter(Soil=="S17", Amendment=="Y") %>% distinct(ASV)
gr.S17n.asv = gr.est %>% filter(Soil=="S17", Amendment=="N") %>% distinct(ASV)

# Overlap
inner_join(gr.C3.asv, gr.S17.asv) %>% nrow() # overall
inner_join(gr.C3y.asv, gr.S17y.asv) %>% nrow() # water-amended 
inner_join(gr.C3n.asv, gr.S17n.asv) %>% nrow() # c-amended
inner_join(gr.C3y.asv, gr.C3n.asv) %>% nrow() # cropped
inner_join(gr.S17y.asv, gr.S17n.asv) %>% nrow() # successional
```


# Growth rate distributions

Average across replicates: makes for cleaner presentation.

```{r}
gr.estavg = gr.est %>% 
  group_by(Soil, Amendment, ASV, Domain, Phylum, Class, Order, Family, Genus) %>% 
  summarize(avgk = mean(k), avgStart = mean(start_day), avgEnd = mean(end_day), avgLength = mean(length_days))
```

Kernel density graphs

See: https://www.statsmodels.org/dev/examples/notebooks/generated/kernel_density.html

```{r, results="show"}
gr.estavg %>% 
  filter(Amendment=="N") %>%
  ggplot(aes(x=log(avgk), color=Soil, fill=Soil)) +
  geom_density(alpha=0.5) +
  labs(title="Only water", x="log specific growth rate", y="Kernel Density") +
  theme_test() +
  theme(#legend.position="none",
        title=element_text(size=12),
        axis.text=element_text(size=12),
        axis.title=element_text(size=12)) +
  scale_color_manual(labels = c("Cropped", "Successional"), values=c("#F8766D","#00BFC4")) +
  scale_fill_manual(labels = c("Cropped", "Successional"), values=c("#F8766D","#00BFC4")) +
  scale_y_continuous(limits=c(0,1)) +
  scale_x_continuous(limits=c(-5,0))

gr.estavg %>% 
  filter(Amendment=="Y") %>%
  ggplot(aes(x=log(avgk), color=Soil, fill=Soil)) +
  geom_density(alpha=0.5) +
  labs(title="C-amended", x="log specific growth rate", y="Kernel Density") +
  theme_test() +
  theme(#legend.position="none",
        title=element_text(size=12),
        axis.text=element_text(size=12),
        axis.title=element_text(size=12)) +
  scale_color_manual(labels = c("Cropped", "Successional"), values=c("#F8766D","#00BFC4")) +
  scale_fill_manual(labels = c("Cropped", "Successional"), values=c("#F8766D","#00BFC4")) +
  scale_y_continuous(limits=c(0,1)) +
  scale_x_continuous(limits=c(-5,0))

gr.estavg %>% 
  filter(Soil=="C3") %>%
  ggplot(aes(x=log(avgk), color=Amendment, fill=Amendment)) +
  geom_density(alpha=0.5) +
  labs(title="Cropped", x="log specific growth rate", y="Kernel Density") +
  theme_test() +
  theme(#legend.position="none",
        title=element_text(size=12),
        axis.text=element_text(size=12),
        axis.title=element_text(size=12)) +
  scale_color_manual(labels = c("C-added", "Water"), values=c("#E78E43", "#423EC5")) +
  scale_fill_manual(labels = c("C-added", "Water"), values=c("#E78E43", "#423EC5")) +
  scale_y_continuous(limits=c(0,1)) +
  scale_x_continuous(limits=c(-5,0))

gr.estavg %>% 
  filter(Soil=="S17") %>%
  ggplot(aes(x=log(avgk), color=Amendment, fill=Amendment)) +
  geom_density(alpha=0.5) +
  labs(title="Successional", x="log specific growth rate", y="Kernel Density") +
  theme_test() +
  theme(#legend.position="none",
        title=element_text(size=12),
        axis.text=element_text(size=12),
        axis.title=element_text(size=12)) +
  scale_color_manual(labels = c("C-added", "Water"), values=c("#E78E43", "#423EC5")) +
  scale_fill_manual(labels = c("C-added", "Water"), values=c("#E78E43", "#423EC5")) +
  scale_y_continuous(limits=c(0,1)) +
  scale_x_continuous(limits=c(-5,0))
```


# Growth rate distribution binning

Hypothesis: growth rate distributions will differ by soil and by amendment.

In order to conduct tests on the distirbution of the data, I need to bin it discretely. The bins will need to be equal width and have the same start and end points across treatments so that they are comparable, and the number of bins must be chosen to simplify the data without losing too much of the "shape" of the distribution (see smoothed histograms above for reference).

Ranging from 5 to 10 bins.

Overall:

```{r}
for (i in 5:10) {
  print(gr.estavg %>%
    ggplot(aes(x=log(avgk))) +
    geom_histogram(bins=i, color="black", fill="gray") +
    theme_test() +
    labs(title=paste("Overall,", i, "bins", sep=" ")))
}
```

Treatments:

```{r}
for (s in c("C3", "S17")) {
  for (a in c("Y", "N")) {
    for (i in 5:10) {
      print(gr.estavg %>%
        filter(Soil==s, Amendment==a) %>%
        ggplot(aes(x=log(avgk))) +
        geom_histogram(bins=i, color="black", fill="gray") +
        theme_test() +
        labs(title=paste(s, a, i, "bins")))
    }
  }
}
```

It looks like 5 bins is too few, 6 or 7 seem to more adequately capture the distributions. I'll choose 7 bins.

There need to be the same number and width of bins in each dataset.

Caculate min, max, and range of specific growth rate values in whole dataset:

```{r, , results="show"}
print(paste("Min:", min(log(gr.estavg$avgk))))
print(paste("Max:", max(log(gr.estavg$avgk))))
print(paste("Range:", max(log(gr.estavg$avgk))-min(log(gr.estavg$avgk))))
```

Calculate bin width based on range:

```{r, results="show"}
width = (max(log(gr.estavg$avgk))-min(log(gr.estavg$avgk)))/7
print(paste("Bin width:", width))

kmin=min(log(gr.estavg$avgk))
print(paste("Bin 1 boundaries: less than", kmin+width))
print(paste("Bin 2 boundaries:", kmin+width, "to",  kmin+width*2))
print(paste("Bin 3 boundaries:", kmin+width*2, "to",  kmin+width*3))
print(paste("Bin 4 boundaries:", kmin+width*3, "to",  kmin+width*4))
print(paste("Bin 5 boundaries:", kmin+width*4, "to",  kmin+width*5))
print(paste("Bin 6 boundaries:", kmin+width*5, "to",  kmin+width*6))
print(paste("Bin 7 boundaries:", kmin+width*6, "and greater"))
```

```{r}
# Bin data based on boundaries chosen and log k
gr.est.bins = gr.est %>% mutate(bin = ifelse(log(k) < kmin+width, 1, ifelse(log(k) < kmin+width*2, 2, ifelse(log(k) < kmin+width*3, 3, ifelse(log(k) < kmin+width*4, 4, ifelse(log(k) < kmin+width*5, 5, ifelse(log(k) < kmin+width*6, 6, 7)))))))

# Calculate bin information (proportion, frequency)
gr.bins = data.frame()
for (s in as.character(unique(gr.est.bins$Soil))) { 
  for (a in as.character(unique(gr.est.bins$Amendment))) {
    for (r in as.character(unique(gr.est.bins$Replicate))) {
      total = nrow(gr.est.bins[gr.est.bins$Soil==s & gr.est.bins$Amendment==a & gr.est.bins$Replicate==r,])
      for (b in as.character(unique(gr.est.bins$bin))) {
        bin = gr.est.bins[gr.est.bins$Soil==s & gr.est.bins$Amendment==a & gr.est.bins$Replicate==r & gr.est.bins$bin==b,]
        binfreq = nrow(bin)
        binratio = binfreq/total
        thisrow = data.frame(s, a, r, b, binfreq, binratio)
        gr.bins = rbind(gr.bins, thisrow)   
      }
    }
  }
}
colnames(gr.bins) = c("Soil","Amendment","Replicate","bin","frequency","proportion")
gr.bins$bin = factor(gr.bins$bin, levels=c(1,2,3,4,5,6,7))
```

Plot bins:

```{r, results="show"}
ggplot(gr.bins) +
  geom_boxplot(aes(x=bin, y=proportion, color=Soil)) +
  facet_wrap(~Amendment) +
  theme_test() +
  labs(x="Slow to fast", y="Proportion of total taxa")

ggplot(gr.bins) +
  geom_boxplot(aes(x=bin, y=frequency, color=Soil)) +
  facet_wrap(~Amendment) +
  theme_test() +
  labs(x="Slow to fast", y="Frequency")

ggplot(gr.bins) +
  geom_boxplot(aes(x=bin, y=proportion, color=Amendment)) +
  facet_wrap(~Soil) +
  scale_color_manual(values=c("#E78E43", "#423EC5")) +
  theme_test() +
  labs(x="Slow to fast", y="Proportion of total taxa")

ggplot(gr.bins) +
  geom_boxplot(aes(x=bin, y=frequency, color=Amendment)) +
  facet_wrap(~Soil) +
  scale_color_manual(values=c("#E78E43", "#423EC5")) +
  theme_test() +
  labs(x="Slow to fast", y="Frequency")
```

**Statistics**

First, I want to know if the distribution of growth rates differs by habitat or amendment. I'll use a contingency table and Fisher's exact test.

See: http://www.biostathandbook.com/fishers.html

```{r}
# Average replicates and round (Frischer's needs integers, not continuous numbers)
gr.bins.avg = gr.bins %>% group_by(Soil, Amendment, bin) %>% summarize(avgfreq = mean(frequency)) %>% mutate(avgfreq.rnd = round(avgfreq))

# Make contingency tables
conting.y = dcast(gr.bins.avg[gr.bins.avg$Amendment=="Y",], Soil~bin, value.var="avgfreq.rnd")
conting.y = conting.y[,-1] %>% as.matrix()

conting.n = dcast(gr.bins.avg[gr.bins.avg$Amendment=="N",], Soil~bin, value.var="avgfreq.rnd")
conting.n = conting.n[,-1] %>% as.matrix()

conting.C3 = dcast(gr.bins.avg[gr.bins.avg$Soil=="C3",], Amendment~bin, value.var="avgfreq.rnd")
conting.C3 = conting.C3[,-1] %>% as.matrix()

conting.S17 = dcast(gr.bins.avg[gr.bins.avg$Soil=="S17",], Amendment~bin, value.var="avgfreq.rnd")
conting.S17 = conting.S17[,-1] %>% as.matrix()

# Run Fisher's exact tests
set.seed(1)
fish.testy = fisher.test(conting.y, simulate.p.value = TRUE)
fish.testn = fisher.test(conting.n, simulate.p.value = TRUE)
fish.testC3 = fisher.test(conting.C3, simulate.p.value = TRUE)
fish.testS17 = fisher.test(conting.S17, simulate.p.value = TRUE)

# Multiple test correction
fish.pvals = c(fish.testn$p.value, fish.testy$p.value, fish.testC3$p.value, fish.testS17$p.value)
fish.padj = p.adjust(fish.pvals, method="holm", n=length(fish.pvals))
print(paste("Unamended cropped vs Succ., adj. p-value:", fish.padj[1]))
print(paste("Amended cropped vs Succ., adj. p-value:", fish.padj[2]))
print(paste("Cropped amended vs unamended, adj. p-value:", fish.padj[3]))
print(paste("Succ. amended vs unamended, adj. p-value:", fish.padj[4]))
```


# Growth rate shift up/down

Among ASVs that were estimated in both cropped and successional soil treatments, do they reliably shift up or down in growth rate?

Hypotheses: 

1. ASVs found in water and C-amended treatments will grow faster in the C-amended treatment in both soils because of higher resource availability.

2. ASVs found in both cropped and successional soils overall will grow faster in the cropped soil because of greater disturbances in the cropped soil.

```{r}
# Isolate treatment data
gr.C3.avgk = gr.est %>% 
  filter(Soil=="C3") %>% 
  group_by(Soil, ASV) %>% 
  summarize(C3.avgk=mean(k))

gr.S17.avgk = gr.est %>%
  filter(Soil=="S17") %>% 
  group_by(Soil, ASV) %>% 
  summarize(S17.avgk=mean(k))

gr.C3y.avgk = gr.est %>% 
  filter(Soil=="C3", Amendment=="Y") %>% 
  group_by(Soil, Amendment, ASV) %>% 
  summarize(C3y.avgk=mean(k))

gr.C3n.avgk = gr.est %>% 
  filter(Soil=="C3", Amendment=="N") %>% 
  group_by(Soil, Amendment, ASV) %>% 
  summarize(C3n.avgk=mean(k))

gr.S17y.avgk = gr.est %>% 
  filter(Soil=="S17", Amendment=="Y") %>% 
  group_by(Soil, Amendment, ASV) %>% 
  summarize(S17y.avgk=mean(k))

gr.S17n.avgk = gr.est %>% 
  filter(Soil=="S17", Amendment=="N") %>% 
  group_by(Soil, Amendment, ASV) %>% 
  summarize(S17n.avgk=mean(k))
```

Visualize:

```{r}
# Overall cropped vs succesional
inner_join(gr.C3.avgk, gr.S17.avgk, by="ASV") %>%
  ggplot(aes(x=C3.avgk, y=S17.avgk)) +
  geom_point() +
  geom_smooth(method="lm") +
  labs(x="Cropped, log specific growth rate", y="Successional, log specific growth rate") +
  theme_test()

# Cropped C-amended vs water
inner_join(gr.C3y.avgk, gr.C3n.avgk, by="ASV") %>% 
  ggplot(aes(x=C3y.avgk, y=C3n.avgk)) +
  geom_point() +
  geom_smooth(method="lm") +
  labs(x="Cropped C-amended, log specific growth rate", y="Cropped water-amended, log specific growth rate") +
  theme_test()

# Succesional C-amended vs water
inner_join(gr.S17y.avgk, gr.S17n.avgk, by="ASV") %>% 
  ggplot(aes(x=S17y.avgk, y=S17n.avgk)) +
  geom_point() +
  geom_smooth(method="lm") +
  labs(x="Successional C-amended, log specific growth rate", y="Successional water-amended, log specific growth rate") +
  theme_test()
```

Doesn't look promising.

**Statistcs**

Try simple linear regression first:

```{r}
# Test overall cropped vs successional using linear regression
gr.C3S17.avgk = inner_join(gr.C3.avgk, gr.S17.avgk, by="ASV")
gr.C3S17.avgk.lm = lm(C3.avgk ~ S17.avgk, data=gr.C3S17.avgk)

# Evaluate assumptions of linear regression
hist(gr.C3S17.avgk.lm$residuals) # evaluate assumption of normality
plot(predict(gr.C3S17.avgk.lm), resid(gr.C3S17.avgk.lm)) # evaluate assumption of homoskedasticity

# P value
summary(gr.C3S17.avgk.lm)$coefficients[2,4]
```

```{r}
# Cropped C-amended vs water
gr.C3yn.avgk = inner_join(gr.C3y.avgk, gr.C3n.avgk, by="ASV")
gr.C3yn.avgk.lm = lm(C3y.avgk ~ C3n.avgk, data=gr.C3yn.avgk)

# Evaluate assumptions of linear regression
hist(gr.C3yn.avgk.lm$residuals) # evaluate assumption of normality
plot(predict(gr.C3yn.avgk.lm), resid(gr.C3yn.avgk.lm)) # evaluate assumption of homoskedasticity

# P value
summary(gr.C3yn.avgk.lm)$coefficients[2,4]
```

```{r}
# Successional C-amended vs water
gr.S17yn.avgk = inner_join(gr.S17y.avgk, gr.S17n.avgk, by="ASV")
gr.S17yn.avgk.lm = lm(S17y.avgk ~ S17n.avgk, data=gr.S17yn.avgk)

# Evaluate assumptions of linear regression
hist(gr.S17yn.avgk.lm$residuals) # evaluate assumption of normality
plot(predict(gr.S17yn.avgk.lm), resid(gr.S17yn.avgk.lm)) # evaluate assumption of homoskedasticity

# P value
summary(gr.S17yn.avgk.lm)$coefficients[2,4]
```

Assumptions look fine, but the successional overlap data is just sparse. Anyways, nothing is significant.


# Explore bimodal distributions - slow and fast groups?

The growth rate distribution graphs show distinct bimodal distributions in the water treatment. This seems to indicate differing populations, which I want to explore further.

Split data into "fast" and "slow" groups based on the bimodal distributions.

**Split by the bimodal distributions in unamended soils:**

Choose thresholds for each soil:

```{r, results="show"}
gr.estavg %>% 
  filter(Soil=="C3") %>%
  ggplot(aes(x=log(avgk), color=Amendment, fill=Amendment)) +
  geom_density(alpha=0.5) +
  theme_test() +
  theme(#legend.position="none",
        title=element_text(size=12),
        axis.text=element_text(size=12),
        axis.title=element_text(size=12)) +
  scale_color_manual(values=c("#E78E43", "#423EC5")) +
  scale_fill_manual(values=c("#E78E43", "#423EC5")) +
  labs(title="Cropped", x="ln speciifc growth rate", y="Density") +
  scale_y_continuous(limits=c(0,1)) +
  scale_x_continuous(limits=c(-5,0)) +
  geom_vline(xintercept=-2.4, linetype=2)

gr.estavg %>% 
  filter(Soil=="S17") %>%
  ggplot(aes(x=log(avgk), color=Amendment, fill=Amendment)) +
  geom_density(alpha=0.5) +
  theme_test() +
  theme(#legend.position="none",
        title=element_text(size=12),
        axis.text=element_text(size=12),
        axis.title=element_text(size=12)) +
  scale_color_manual(values=c("#E78E43", "#423EC5")) +
  scale_fill_manual(values=c("#E78E43", "#423EC5")) +
  labs(title="Successional", x="ln speciifc growth rate", y="Density") +
  scale_y_continuous(limits=c(0,1)) +
  scale_x_continuous(limits=c(-5,0)) +
  geom_vline(xintercept=-1.9, linetype=2)
```

Label ASVs based on avg growth rate threshold:

```{r}
# Water only
gr.estavg.group = gr.estavg %>% 
  filter(Amendment=="N") %>% 
  mutate(group = ifelse(Soil=="C3" & log(avgk) < -2.4, "slow", ifelse(Soil=="S17" & log(avgk) < -1.9, "slow", "fast")))

gr.estavg.group = gr.estavg.group[,c(1,3:ncol(gr.estavg.group))]
```

```{r, eval=FALSE}
# save 
saveRDS(gr.estavg.group, file="rdata.files/gr.groups.rds")
```


### "Fast" vs "slow in unamended soil

Is there  a difference in the proportion of these groups in the cropped and successional water-amended soil?

Label individual replicates

```{r}
# Label taxa in replicates based on thresholds chosen from distributions
gr.est.group = gr.est %>% 
  filter(Amendment=="N") %>% # water only treatment
  mutate(group = ifelse(Soil=="C3" & log(k) < -2.4, "slow", ifelse(Soil=="S17" & log(k) < -1.9, "slow", "fast")))

# Number of taxa in each group per rep
gr.est.group.count = gr.est.group %>%
  group_by(Soil, Amendment, Replicate, group) %>%
  summarize(n = n())

# Proportion fast vs slow in each rep
gr.est.group.prop = gr.est.group.count %>%
  pivot_wider(names_from="group", values_from="n") %>%
  mutate(prop_fastslow = slow/(fast+slow))
```

**Visualize**

```{r}
gr.est.group.count %>%
  ggplot(aes(x=group, y=n, color=Replicate)) +
  geom_point() +
  facet_wrap(~Soil) +
  theme_test()

gr.est.group.prop %>%
  ggplot(aes(x=Soil, y=prop_fastslow)) +
  geom_point() +
  theme_test()
```

**Statistics**

Compare number of taxa in "fast" vs "slow" group in each soil.

Linear mixed-effects model with replicate fitted as random to account for repeated measures from replicates (slow and fast).

```{r}
group.count.C3.lmer = lmer(log(n) ~ group + (1|Replicate), data=gr.est.group.count[gr.est.group.count$Soil=="C3",])
hist(residuals(group.count.C3.lmer))
plot(residuals(group.count.C3.lmer), predict(group.count.C3.lmer))
summary(group.count.C3.lmer)

group.count.S173.lmer = lmer(log(n) ~ group + (1|Replicate), data=gr.est.group.count[gr.est.group.count$Soil=="S17",])
hist(residuals(group.count.S173.lmer))
plot(residuals(group.count.S173.lmer), predict(group.count.S173.lmer))
summary(group.count.S173.lmer)
```

Test proportion of fast to slow taxa in cropped and successional soil (another way to test)

Using simple linear regression

```{r}
group.prop.lm = lm(log(prop_fastslow) ~ Soil, data=gr.est.group.prop)
hist(residuals(group.prop.lm))
plot(residuals(group.prop.lm), predict(group.prop.lm))
summary(group.prop.lm)
```

Not significant. Likely not enough power.


### Abundances of groups in field communities

Do these taxa exist in the in-field community data? If so, what are they're abundances? I.e. did the soil shape the growth dynamics we observed in the microcosms?

Hypothesis: abundances of slow and fast taxa within field communities differ between cropped and successional soils.

Import field data:

```{r}
# Import data
field = readRDS("rdata.files/field.norm.cleaned.rds")

# Add group labels
field.group = field %>% 
  inner_join(gr.estavg.group, var=c("Soil", "ASV", "Domain", "Phylum", "Class", "Order", "Family", "Genus",)) %>%
  filter(!norm_abund==0) %>% # remove 0 values, ie ASV not present
  mutate(group = parse_factor(group, levels=c("slow", "fast"))) # more intuitive order for groups
```


#### Individual ASV normalized abundances in each group:

```{r}
# Visualize
field.group %>% 
  ggplot(aes(x=group, y=log(norm_abund), color=as.factor(Block))) +
  geom_boxplot() +
  geom_jitter(alpha=0.2) +
  facet_wrap(~Soil) +
  labs(title="Individual ASV abundances", x="", y="natural log internal-standard normalized abundance", color="Block") +
  theme_test()
```

#### Total, aggregated abundances in each group:

```{r}
# Sum abundances in each group
field.group.sum = field.group %>% 
  group_by(Soil, Block, Replicate, group) %>%
  summarize(total_abund = sum(norm_abund))

# Visualize
field.group.sum %>% 
  ggplot(aes(x=group, y=log(total_abund), color=as.factor(Block))) +
  geom_boxplot() +
  geom_jitter(alpha=0.2) +
  facet_wrap(~Soil) +
  labs(title="Total, aggregated abundance", x="", y="natural log internal-standard normalized abundance", color="Block") +
  theme_test()

field.group.sum %>% 
  ggplot(aes(x=group, y=log(total_abund))) +
  geom_boxplot() +
  geom_jitter(alpha=0.2) +
  facet_wrap(~Soil) +
  labs(title="Total, aggregated abundance", x="", y="natural log internal-standard normalized abundance", color="Block") +
  theme_test()
```

Strong block effect in cropped soil.


# Figures

### Distribution of growth rates

```{r}
p1 = gr.estavg %>%
  mutate(Soil = recode_factor(Soil, S17="Successional", C3="Cropped"),
         Amendment = recode_factor(Amendment, N="Water", Y="C-added")) %>%
  ggplot(aes(x=log(avgk), fill=Soil)) +
  geom_density(alpha=0.5) +
  facet_wrap(~Amendment, ncol=1) +
  scale_fill_manual(values=c(NA, "gray")) +
  labs(y="Kernel density", x="Specific growth rate") +
  theme_test() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        strip.background = element_blank(),
        legend.position = c(.25,.93),
        legend.text = element_text(size=8),
        legend.key.width=unit(0.5, "cm"),
        legend.key.height=unit(0.5, "cm"),
        legend.title = element_blank(),
        legend.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.title.x = element_text(margin = unit(c(1, 0, 0, 0), "mm"), size=10),
        axis.title.y = element_text(margin = unit(c(0, 1.5, 0, 0), "mm"), size=10),
        axis.text = element_text(size = 8))
p1

p2 = gr.bins %>%
  mutate(Soil = recode_factor(Soil, S17="Successional", C3="Cropped"),
         Amendment = recode_factor(Amendment, N="Water", Y="C-added")) %>%
  ggplot(aes(x=bin, y=frequency, fill=Soil)) +
  geom_boxplot(alpha=0.5) +
  facet_wrap(~Amendment, ncol=1) +
  scale_fill_manual(values=c(NA, "gray")) +
  labs(x="Bin number", y="Number of ASVs") +
  theme_test() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        strip.background = element_blank(),
        legend.position = "none",
        axis.line = element_line(colour = "black"),
        axis.title.x = element_text(margin = unit(c(1, 0, 0, 0), "mm"), size=10),
        axis.title.y = element_text(margin = unit(c(0, 1.5, 0, 0), "mm"), size=10),
        axis.text = element_text(size = 8))
p2

habitatdist.plot = plot_grid(p1, NULL, p2, nrow=1, align="h", rel_widths=c(1,0.04,1), labels=c("a", NA, "b"), label_size=10)
habitatdist.plot
```

```{r, eval=FALSE}
ggsave(plot=habitatdist.plot, file="figures/fig_habitatdist.pdf", device=pdf, width=140, height=90, units="mm")
```

### Shift in growth rate

```{r}
p3 = inner_join(gr.C3y.avgk, gr.C3n.avgk, by="ASV") %>% 
  ggplot(aes(x=log(C3y.avgk), y=log(C3n.avgk))) +
  geom_point(shape=1) +
  #geom_smooth(method="lm", color="black", linetype=2) +
  labs(x="Cropped, C-amended", y="Cropped, water") +
  scale_x_continuous(limits=c(-3,0)) +
  scale_y_continuous(limits=c(-3,0)) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        strip.background = element_blank(),
        legend.position = "none",
        axis.line = element_line(colour = "black"),
        axis.title.x = element_text(margin = unit(c(1, 0, 0, 0), "mm"), size=10),
        axis.title.y = element_text(margin = unit(c(0, 1.5, 0, 0), "mm"), size=10),
        axis.text = element_text(size = 8))
p3

p4 = inner_join(gr.S17y.avgk, gr.S17n.avgk, by="ASV") %>% 
  ggplot(aes(x=log(S17y.avgk), y=log(S17n.avgk))) +
  geom_point(shape=1) +
  #geom_smooth(method="lm", color="black", linetype=2) +
  labs(x="Successional, C-amended", y="Succesional, water") +
  scale_x_continuous(limits=c(-3,0)) +
  scale_y_continuous(limits=c(-3,0)) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        strip.background = element_blank(),
        legend.position = "none",
        axis.line = element_line(colour = "black"),
        axis.title.x = element_text(margin = unit(c(1, 0, 0, 0), "mm"), size=10),
        axis.title.y = element_text(margin = unit(c(0, 1.5, 0, 0), "mm"), size=10),
        axis.text = element_text(size = 8))
p4

p5 = inner_join(gr.C3y.avgk, gr.S17y.avgk, by="ASV") %>% 
  ggplot(aes(x=log(C3y.avgk), y=log(S17y.avgk))) +
  geom_point(shape=1) +
  #geom_smooth(method="lm", color="black", linetype=2) +
  labs(x="Cropped, C-amended", y="Succesional, C-amended") +
  scale_x_continuous(limits=c(-3,0)) +
  scale_y_continuous(limits=c(-3,0)) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        strip.background = element_blank(),
        legend.position = "none",
        axis.line = element_line(colour = "black"),
        axis.title.x = element_text(margin = unit(c(1, 0, 0, 0), "mm"), size=10),
        axis.title.y = element_text(margin = unit(c(0, 1.5, 0, 0), "mm"), size=10),
        axis.text = element_text(size = 8))
p5

p6 = inner_join(gr.C3n.avgk, gr.S17n.avgk, by="ASV") %>% 
  ggplot(aes(x=log(C3n.avgk), y=log(S17n.avgk))) +
  geom_point(shape=1) +
  #geom_smooth(method="lm", color="black", linetype=2) +
  labs(x="Cropped, water", y="Succesional, water") +
  scale_x_continuous(limits=c(-3,0)) +
  scale_y_continuous(limits=c(-3,0)) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        strip.background = element_blank(),
        legend.position = "none",
        axis.line = element_line(colour = "black"),
        axis.title.x = element_text(margin = unit(c(1, 0, 0, 0), "mm"), size=10),
        axis.title.y = element_text(margin = unit(c(0, 1.5, 0, 0), "mm"), size=10),
        axis.text = element_text(size = 8))
p6

shift.plot = plot_grid(p3, p4, p5, p6)
shift.plot

```

```{r, eval=FALSE}
ggsave(shift.plot, file="figures/suppfig_shiftgr.pdf", device=pdf, width=140, height=140, units="mm")
```

### Start and end of growth

Note sure if will include.

```{r}
p7 = gr.est.avg %>%
  mutate(Soil = recode_factor(Soil, C3="Cropped", S17="Successional"),
         Amendment = recode_factor(Amendment, Y="C-amended", N="Water")) %>%
  ggplot(aes(x=Soil, y=avg_start)) +
  geom_boxplot(aes(fill=Amendment)) +
  scale_fill_manual(values=c(NA, "gray")) +
  scale_y_continuous(limits=c(0,4)) +
  labs(title="Start of growth", y="Day") +
  theme(legend.title = element_blank(),
        legend.key = element_blank(),
        legend.position = c(0.15, 0.8),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        strip.background = element_blank(),
        title = element_text(size = 10),
        axis.title.x = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text = element_text(size = 8))
p7

p8 = gr.est.avg %>%
  mutate(Soil = recode_factor(Soil, C3="Cropped", S17="Successional"),
         Amendment = recode_factor(Amendment, Y="C-amended", N="Water")) %>%
  ggplot(aes(x=Soil, y=avg_end)) +
  geom_boxplot(aes(fill=Amendment)) +
  scale_fill_manual(values=c(NA, "gray")) +
  scale_y_continuous(limits=c(0,10)) +
  labs(title="End of growth", y="Day") +
  theme(legend.position = "none",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        strip.background = element_blank(),
        title = element_text(size = 10),
        axis.title.x = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text = element_text(size = 8))
p8

startend.plots = plot_grid(p7, p8, nrow=2, labels=c("a", "b"), label_size=10)
startend.plots
```

```{r, eval=FALSE}
ggsave(startend.plots, file="figures/fig_startend.pdf", device=pdf, width=140, height=90, units="mm")
```

