---
title: "Dunlop growth rates - 16S copy number"
author: "Cassandra Wattenburger"
date: "11/11/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(results = "show")
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

**Goal of this script:** Use PAPRICA results to explore how in-situ growth rate relates to predicted genomic features of bacteria.

Notes:

* 16S copy number estimations were conducted using PAPRICA v0.5 (vs 0.6 had placement issues)
* Only representative sequences of ASVS with growth rate estimates were used for PAPRICA analysis

# Import libraries

```{r}
library(tidyverse)
library(cowplot)
library(MuMIn)

sessionInfo()

rm(list=ls())
```

# Import PAPRICA and growth estimation data

* PAPRICA results
* in-situ growth rate estimates (cleaned)

```{r}
# Import PAPRICA output
papEdge = read_csv("copynum_est/paprica0.5_results/growthrate.seqs.subset.bacteria.edge_data.csv")
papSeqs = read_csv("copynum_est/paprica0.5_results/growthrate.seqs.subset.bacteria.combined_16S.bacteria.tax.clean.unique.align.csv")

# Merge and remove unnecessary columns
papEdge2 = papEdge %>% select(edge_num = X1, nedge, genome_size, n16S)
papSeqs2 = papSeqs %>% select(edge_num, ASV=name, distal_length, pendant_length)
paprica = merge(papEdge2, papSeqs2, by="edge_num") %>% select(ASV, edge_num, nedge, distal_length, pendant_length, genome_size, n16S)
paprica

# Import growth rate estimates
grest = readRDS("rdata.files/gr_gr.final_rm0.rds") %>% select(ASV, Soil, Amendment, Replicate, k)

# Import taxonomic info
tax = read_tsv("output/growthrate.taxonomy.final.tsv") %>% select(ASV=OTU, everything())

# Merge PAPRICA results with growth rate estimates and taxonomy
paprica.gr = inner_join(paprica, grest, by="ASV")
paprica.gr = inner_join(paprica.gr, tax, by="ASV")

# Average growth rates across replicates
paprica.gr.avgall = paprica.gr %>% # overall
  group_by(ASV, Phylum, Class, Order, Family, Genus) %>%
  summarize(avgk=mean(k), n16S=mean(n16S), genome_size=mean(genome_size), distal_length=mean(distal_length))

paprica.gr.avgsoil = paprica.gr %>% # by soil
  group_by(ASV, Soil, Phylum, Class, Order, Family, Genus) %>%
  summarize(avgk=mean(k), n16S=mean(n16S), genome_size=mean(genome_size), distal_length=mean(distal_length))

paprica.gr.avgtrt = paprica.gr %>% # by treatment 
  group_by(ASV, Soil, Amendment, Phylum, Class, Order, Family, Genus) %>% 
  summarize(avgk=mean(k), n16S=mean(n16S), genome_size=mean(genome_size), distal_length=mean(distal_length)) # PAPRICA results identical b/w replicates


```

# Average distal lengths

The distal length is the phylogenetic distance from the edge where each ASV was placed on the reference tree to the furthest tip on the tree descending from that edge. We want this not to exceed 15% for the PAPRICA results to be reliable, based on results from Louca et al. 2018.

Louca, S., Doebeli, M. & Parfrey, L. W. Correcting for 16S rRNA gene copy numbers in microbiome surveys remains an unsolved problem. Microbiome 6, 41 (2018).

Numbers listed as mean, min, max, standard deviation.

Overall

```{r}
# Overall
mean(paprica.gr.avgall$distal_length)
min(paprica.gr.avgall$distal_length)
max(paprica.gr.avgall$distal_length)
sd(paprica.gr.avgall$distal_length)
```

Cropped
```{r}
# Cropped
mean(paprica.gr.avgsoil[paprica.gr.avgsoil$Soil=="C3",]$distal_length)
min(paprica.gr.avgsoil[paprica.gr.avgsoil$Soil=="C3",]$distal_length)
max(paprica.gr.avgsoil[paprica.gr.avgsoil$Soil=="C3",]$distal_length)
sd(paprica.gr.avgsoil[paprica.gr.avgsoil$Soil=="C3",]$distal_length)
```

Successional
```{r}
# Successional
mean(paprica.gr.avgsoil[paprica.gr.avgsoil$Soil=="S17",]$distal_length)
min(paprica.gr.avgsoil[paprica.gr.avgsoil$Soil=="S17",]$distal_length)
max(paprica.gr.avgsoil[paprica.gr.avgsoil$Soil=="S17",]$distal_length)
sd(paprica.gr.avgsoil[paprica.gr.avgsoil$Soil=="S17",]$distal_length)
```

Cropped, C-amended
```{r}
# Cropped, C-amended
mean(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="C3" & paprica.gr.avgtrt$Amendment=="Y",]$distal_length)
min(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="C3" & paprica.gr.avgtrt$Amendment=="Y",]$distal_length)
max(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="C3" & paprica.gr.avgtrt$Amendment=="Y",]$distal_length)
sd(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="C3" & paprica.gr.avgtrt$Amendment=="Y",]$distal_length)
```

Cropped, water
```{r}
# Cropped, water
mean(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="C3" & paprica.gr.avgtrt$Amendment=="N",]$distal_length)
min(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="C3" & paprica.gr.avgtrt$Amendment=="N",]$distal_length)
max(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="C3" & paprica.gr.avgtrt$Amendment=="N",]$distal_length)
sd(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="C3" & paprica.gr.avgtrt$Amendment=="N",]$distal_length)
```

Successional, C-amended
```{r}
# Successional, C-amended
mean(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="S17" & paprica.gr.avgtrt$Amendment=="Y",]$distal_length)
min(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="S17" & paprica.gr.avgtrt$Amendment=="Y",]$distal_length)
max(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="S17" & paprica.gr.avgtrt$Amendment=="Y",]$distal_length)
sd(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="S17" & paprica.gr.avgtrt$Amendment=="Y",]$distal_length)
```

Successional, water
```{r}
# Successional, water
mean(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="S17" & paprica.gr.avgtrt$Amendment=="N",]$distal_length)
min(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="S17" & paprica.gr.avgtrt$Amendment=="N",]$distal_length)
max(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="S17" & paprica.gr.avgtrt$Amendment=="N",]$distal_length)
sd(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="S17" & paprica.gr.avgtrt$Amendment=="N",]$distal_length)
```

High variation, but these distances look good. There may be ASVs that are much more distantly related to the reference genomes than the average, but for the most part these numbers suggest that the ASVs are well-represented by the reference genomes. Additionally, this is the distance to the furthest neighbor, not the nearest, so these numbers are over-inflated to begin with.

Interestingly, the C-amended treatments typically have lower average distal lengths than the water-amended treatments. Reference genomes had to be cultured and isolated to be sequenced and it makes sense that more "culturable" ASVs might grow in a soil treatment that was amended with simple sugars.


# Relationship between 16S copy number and in-situ growth rates

Main Hypothesis: positive relationship between in-situ growth rate and 16S copy number (more copies = faster maximal growth)

Null Hypothesis: no relationship between in-situ growth rate and 16S copy number (in-situ != maximal growth, stochasticity in growth response adds noise to grwoth rate estimations)

Visualize:

```{r}
# Visualize
paprica.gr.avgall %>% ggplot(aes(x=avgk, y=n16S)) +
  geom_point(aes(color=Phylum)) +
  geom_smooth(method="lm") +
  labs(x="avg specific growth rate", y="estimated 16S copies") +
  theme_test()

paprica.gr.avgsoil %>% ggplot(aes(x=avgk, y=n16S)) +
  geom_point(aes(color=Phylum)) +
  geom_smooth(method="lm") +
  facet_wrap(~Soil) +
  labs(x="avg specific growth rate", y="estimated 16S copies") +
  theme_test()

paprica.gr.avgtrt %>% ggplot(aes(x=avgk, y=n16S)) +
  geom_point(aes(color=Phylum)) +
  geom_smooth(method="lm") +
  facet_wrap(Soil~Amendment) +
  labs(x="avg specific growth rate", y="estimated 16S copies") +
  theme_test()
```

Colors are phyla.

**Statistics:**

Try simple linear regression first.

```{r}
# Linear regression on log-transformed growth rate
n16S.S17y.lm = lm(avgk~n16S, paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="S17" & paprica.gr.avgtrt$Amendment=="Y",])

# Check assumptions
hist(n16S.S17y.lm$residuals) # assumption of normality
plot(predict(n16S.S17y.lm), resid(n16S.S17y.lm)) # assumption of homoskedasticity
```

Normality looks fine, but there is heterskedasticity and there are outliers. 

Instead, Spearman correlation with permutation, because the data inherently contains ties. 
See: https://stats.stackexchange.com/questions/50015/spearman-correlation-in-the-presence-of-many-ties-how-to-spot-a-problem

```{r}
# Create permuted spearman correlation function
spearman_permute = function(x, y) {
  set.seed(1) # make sure it is reproducible
  perm.estimates = c()
  for (i in 1:1000) { # permute 1000 times
    perm = sample(y) # randomize response variable
    perm.test = suppressWarnings(cor.test(x, perm, method="spearman", alternative="greater")) # spearman correlation with permuted data
    perm.estimates = append(perm.estimates, perm.test$estimate) # permuted estimates
  }
  actual.test = suppressWarnings(cor.test(x, y, method="spearman", alternative="greater")) # spearman correlation with actual data
  actual.estimate = actual.test$estimate # actual data estimate (rho)
  est.exceed = sum(abs(perm.estimates) > actual.estimate) # number of permutations that exceed the non-permuted data rho
  est.equal = sum(abs(perm.estimates) == actual.estimate) # number of permutations that equal the non-permuted data rho
  permuted.pval = (est.exceed + est.equal/2) / length(perm.estimates) # proportion of exceedances: the p-value
  return(c(permuted.pval, actual.estimate)) # the p=value (expand out to include actual estimate)
}

# Tests
n16S.C3y.spearperm = spearman_permute(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="C3" & paprica.gr.avgtrt$Amendment=="Y",]$n16S, paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="C3" & paprica.gr.avgtrt$Amendment=="Y",]$avgk)

n16S.C3n.spearperm = spearman_permute(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="C3" & paprica.gr.avgtrt$Amendment=="N",]$n16S, paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="C3" & paprica.gr.avgtrt$Amendment=="N",]$avgk)

n16S.S17y.spearperm = spearman_permute(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="S17" & paprica.gr.avgtrt$Amendment=="Y",]$n16S, paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="S17" & paprica.gr.avgtrt$Amendment=="Y",]$avgk)

n16S.S17n.spearperm = spearman_permute(paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="S17" & paprica.gr.avgtrt$Amendment=="N",]$n16S, paprica.gr.avgtrt[paprica.gr.avgtrt$Soil=="S17" & paprica.gr.avgtrt$Amendment=="N",]$avgk)

# Multiple test correction
n16S.spearperm.pvals = c(n16S.C3y.spearperm[1], n16S.C3n.spearperm[1], n16S.S17y.spearperm[1], n16S.S17n.spearperm[1])
n16S.spearperm.padj = p.adjust(n16S.spearperm.pvals, method="hochberg", n=length(n16S.spearperm.pvals))

# Estimates
n16S.spearperm.rhos = c(n16S.C3y.spearperm[2], n16S.C3n.spearperm[2], n16S.S17y.spearperm[2], n16S.S17n.spearperm[2])

# Results table
n16S.spearperm.results = data.frame(Soil = c(rep("C3", 2), rep("S17", 2)), Amendment=c("Y","N","Y","N"), p.adj=n16S.spearperm.padj, rho = n16S.spearperm.rhos)
n16S.spearperm.results
```

There are week positive correlations for most treatments except cropped water.


# Explore bimodal distributions

I want to make sure that the trends found in the other script hold up when the data is normalized by the 16S copy number, which is a confounding factor since faster-growing taxa tend to have higher 16S copy number than slower growing taxa, which would inflate the read count abundances from sequencing.

Import field data:

```{r}
# Import field community data
field = readRDS("rdata.files/field.norm.cleaned.rds")

# Import grouped ASVs
gr.group = readRDS("rdata.files/gr.groups.rds")

# Add group labels
field.group = field %>% 
  inner_join(gr.group, var=c("Soil", "ASV", "Domain", "Phylum", "Class", "Order", "Family", "Genus",)) %>%
  filter(!norm_abund==0) %>% # remove 0 values, ie ASV not present
  mutate(group = parse_factor(group, levels=c("slow", "fast"))) # more intuitive order for groups
```

Correct normalized abundances for 16S copy number (divide by n16S)

```{r}
# Add n16S predictions and corect abundances
field.group.n16S = field.group %>% 
  inner_join(paprica, var="ASV") %>%
  mutate(corr_abund = norm_abund/n16S)
```

### Individual ASV abundances

```{r}
# Visualize
field.group.n16S %>% 
  ggplot(aes(x=group, y=log(corr_abund), color=as.factor(Block))) +
  geom_boxplot() +
  geom_jitter(alpha=0.2) +
  facet_wrap(~Soil) +
  labs(title="Individual ASV abundances", x="", y="natural log internal-standard normalized abundance", color="Block") +
  theme_test()
```

**Statistics**

REVISIT MODELS, CHANGES SIGNIFICANCE, SEE WHICH MODEL IS BETTER FIT AND GO WITH THAT

Linear-mixed effects model

* Block as random effect
* Repeated measures because slow and fast are measured from same plot for each replicaterom same sample).

I'm going to try two different models, one that has an interaction effect with block, because I think that may be the case.

Cropped soil:

```{r}
# Cropped
field.group.ind.C3.lmer1 = lmer(log(corr_abund) ~ group + Block + (1|Replicate), data=field.group.n16S[field.group.n16S$Soil=="C3",]) # fit model 1

field.group.ind.C3.lmer2 = lmer(log(corr_abund) ~ group * Block + (1|Replicate), data=field.group.n16S[field.group.n16S$Soil=="C3",]) # fit model 2

# Compare models
# Check r-squared of each
r.squaredGLMM(field.group.ind.C3.lmer1)
r.squaredGLMM(field.group.ind.C3.lmer2)
# Check results
summary(field.group.ind.C3.lmer1)
summary(field.group.ind.C3.lmer2)
```

Not sure what to think. One model says that group is highly significant, the other insignificant. They both explain about the same amount of variance. That's fishy. Proceed with caution.

Successional soil:

```{r}
# Successional
field.group.ind.S17.lmer1 = lmer(log(corr_abund) ~ group + Block + (1|Replicate), data=field.group.n16S[field.group.n16S$Soil=="S17",]) # fit model 1

field.group.ind.S17.lmer2 = lmer(log(corr_abund) ~ group * Block + (1|Replicate), data=field.group.n16S[field.group.n16S$Soil=="S17",]) # fit model 2

# Compare models
# Check r-squared of each
r.squaredGLMM(field.group.ind.S17.lmer1)
r.squaredGLMM(field.group.ind.S17.lmer2)
# Check results
summary(field.group.ind.S17.lmer1)
summary(field.group.ind.S17.lmer2)
```

Same situation, but at least the model results agree with one another.

Model 1 results:

```{r}
plot(resid(field.group.ind.S17.lmer1)) # check normality
plot(predict(field.group.ind.S17.lmer1), resid(field.group.ind.S17.lmer1)) # check variances
summary(field.group.ind.S17.lmer1) # results
```


### Total abundances

```{r}
# Calculate total aggregated abundances
field.group.n16S.total = field.group.n16S %>% 
  group_by(Soil, Block, Replicate, group) %>%
  summarize(total_abund = sum(norm_abund))

# Visualize
field.group.n16S.total %>% 
  ggplot(aes(x=group, y=log(total_abund), color=as.factor(Block))) +
  geom_boxplot() +
  geom_jitter(alpha=0.2) +
  facet_wrap(~Soil) +
  labs(title="Total, aggregated abundance", x="", y="natural log internal-standard normalized abundance", color="Block") +
  theme_test()

field.group.n16S.total %>% 
  ggplot(aes(x=group, y=log(total_abund))) +
  geom_boxplot() +
  geom_jitter(alpha=0.2) +
  facet_wrap(~Soil) +
  labs(title="Total, aggregated abundance", x="", y="natural log internal-standard normalized abundance", color="Block") +
  theme_test()
```

**Statistics**

Linear-mixed effects model.

* Block as random effect
* Repeated measures because slow and fast are measured from same plot for each replicate

```{r}
# Cropped
field.group.sum.C3.lmer = lmer(log(total_abund) ~ group + Block + (1|Replicate), data=field.group.n16S.total[field.group.n16S.total$Soil=="C3",]) # fit model
plot(resid(field.group.sum.C3.lmer)) # check normality
plot(predict(field.group.sum.C3.lmer), resid(field.group.sum.C3.lmer)) # check variances
anova(field.group.sum.C3.lmer) # results

# Successional
field.group.sum.S17.lmer = lmer(log(total_abund) ~ group + Block + (1|Replicate), data=field.group.n16S.total[field.group.n16S.total$Soil=="S17",]) # fit model
plot(resid(field.group.sum.S17.lmer)) # check normality
plot(predict(field.group.sum.S17.lmer), resid(field.group.sum.S17.lmer)) # check variances
anova(field.group.sum.S17.lmer) # results
```

No change in results. Trends were not confounded by 16S copy number.


# Figures

### 16S copy number vs in-situ growth rate

```{r}
# Overall
p1 = paprica.gr.avgtrt %>%
  mutate(Soil = fct_relevel(Soil, "S17", "C3"), 
         Soil = recode_factor(Soil, S17="Successional", C3="Cropped"),
         Amendment = recode_factor(Amendment, Y="C-added", N="Water")) %>%
  ggplot(aes(x=log(avgk), y=n16S)) +
  geom_point(shape=1, alpha=0.6) +
  geom_smooth(method="lm", linetype=2, color="black", alpha=0.8, fill="gray") +
  labs(title="Overall", x=expression("Specific growth rate (ln day"^{-1}*")"), y="Estimated 16S copy number") +
  theme_test() +
  theme(axis.line = element_line(colour = "black"),
        strip.background = element_blank(),
        strip.text = element_text(size=10),
        axis.title.x = element_text(margin = unit(c(1, 0, 0, 0), "mm"), size=10),
        axis.title.y = element_text(margin = unit(c(0, 1, 0, 0), "mm"), size=10))
p1

# By treatment
p2 = paprica.gr.avgtrt %>%
  mutate(Soil = fct_relevel(Soil, "S17", "C3"), 
         Soil = recode_factor(Soil, S17="Successional", C3="Cropped"),
         Amendment = recode_factor(Amendment, Y="C-added", N="Water")) %>%
  ggplot(aes(x=log(avgk), y=n16S)) +
  geom_point(shape=1, alpha=0.6) +
  geom_smooth(method="lm", linetype=2, color="black", alpha=0.8, fill="gray") +
  facet_grid(Amendment ~ Soil, switch="y") +
  labs(x=expression("Specific growth rate (ln day"^{-1}*")")) +
  theme_test() +
  theme(axis.line = element_line(colour = "black"),
        strip.background = element_blank(),
        strip.text = element_text(size=10),
        axis.title.x = element_text(margin = unit(c(1, 0, 0, 0), "mm"), size=10),
        axis.title.y = element_blank())
p2

copynum.plot = plot_grid(p1, NULL, p2, nrow=1, axis="t", align="h", rel_widths=c(1,0.02,1), labels=c("a", NA, "b"), label_size=10)
copynum.plot
```

I'll make small aesthetic tweaks to this oustide of ggplot2, probably.

* make font sizes equal
* make sure plots are aligned vertically

```{r, eval=FALSE}
ggsave(plot=copynum.plot, file="figures/fig_copynum.pdf", device=pdf, width=190, height=100, unit="mm")
```


### Group abudances, corrected for copy number

```{r}
ASVlabs = list()

p3 = field.group.n16S.total %>%
  mutate(Soil = fct_relevel(Soil, "S17"), 
         Soil = recode_factor(Soil, S17="Successional", C3="Cropped")) %>%
  ggplot(aes(x=group, y=log(total_abund))) +
  geom_boxplot() +
  #geom_text() +
  facet_wrap(~Soil) +
  labs(y="Total normalized abundance") +
  theme_test() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        strip.background = element_blank(),
        axis.title.x = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text = element_text(size = 8))
p3
```

```{r, eval=FALSE}
ggsave(plot=p3, file="figures/fig_totalabund.pdf", device=pdf, width=90, height=70, units="mm")
```

