---
title: "Dunlop growth rate - miscellanious"
author: "Cassandra Wattenburger"
date: "11/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(results = "show")
```

**Goal of this script:** Further visualizations and analyses that don't fit into a particular theme or category.


# Import libraries

```{r}
library(tidyverse)
library(ape)
library(phyloseq)

set.seed(4)

sessionInfo()

rm(list=ls())
```


# Import data

```{r}
ucosm = readRDS("rdata.files/ucosm.norm.cleaned.rds")
field = readRDS("rdata.files/field.norm.cleaned.rds")
gr.est = readRDS("rdata.files/gr.final.cleaned.rds")
```

# DNA conc. over time

```{r}
# Import DNA conc. info
dna = read_csv("growthrate_dnaconc.csv") %>% select("Sample" = SampleID, "dna_conc" = Conc.)
ucosm.dna = inner_join(ucosm, dna, by="Sample")

ucosm.dna %>% 
  group_by(Soil, Amendment, Day) %>%
  summarize(dna_conc = mean(dna_conc)) %>%
  mutate(Amendment = recode_factor(Amendment, Y="C-added", N="Water"),
         Soil = recode_factor(Soil, S17="Successional", C3="Cropped")) %>%
  ggplot(aes(x=Day, y=dna_conc, linetype=Soil)) +
  facet_wrap(~Amendment) +
  geom_point() +
  geom_line() +
  labs(y="DNA concentration (ng/μL)") +
  theme_test() +
  theme(strip.background = element_blank(),
        legend.title = element_blank())
```

There appears to be a decrease in DNA concentration over time, particularly in the C-added soil. Not what I expected. Perhaps priming effec of some sort?

# Beta diversity

Import data into phyloseq:

```{r}
# Merge data from microcosm and field

# Add sample type labels
ucosm = mutate(ucosm, Type = "Microcosm")
field = mutate(field, Type = "Field")
ucosm.field = bind_rows(ucosm, field)

# ASV table
ucosm.field.count = ucosm.field %>%
  select(Sample, norm_abund, ASV) %>%
  pivot_wider(names_from = ASV, values_from = norm_abund) %>%
  column_to_rownames(var="Sample")
ucosm.field.count = t(ucosm.field.count)
ucosm.field.count = as.matrix(ucosm.field.count)

# Taxonomy
# GIVING ERROR, NOT SURE WHY
#ucosm.field.tax = ucosm.field %>%
#  select(ASV, Domain, Phylum, Class, Order, Family, Genus) %>%
#  unique() %>%
#  as.data.frame()
#rownames(ucosm.field.tax) = ucosm.field.tax$ASV
#ucosm.field.tax = ucosm.field.tax[,-1]
#ucosm.field.tax = as.matrix(ucosm.field.tax)

# Metadata
ucosm.field.meta = ucosm.field %>%
  select(Sample, Soil, Amendment, Replicate, Day, Block, Type) %>%
  unique() %>%
  mutate(Soil = recode_factor(Soil, S17="Successional", C3="Cropped"),
         Amendment = recode_factor(Amendment, Y="C-amended", N="Water"))
rownames(ucosm.field.meta) = ucosm.field.meta$Sample

# Read phylogenetic tree
tree = read.tree("output/growthrate.tree.final.nwk")

# Create phyloseq object
physeq = phyloseq(otu_table(ucosm.field.count, taxa_are_rows = TRUE),
                  #tax_table(ucosm.field.tax),
                  sample_data(ucosm.field.meta, errorIfNULL = TRUE),
                  tree)
```

### Field vs ucosm

How much did processing the soil and placing it in microcosms change the composition? Hopefully not much.

Ordinate:
```{r}
physeq.day0 = subset_samples(physeq, Day %in% c(NA, 0))
day0.ord = ordinate(physeq.day0, method="NMDS", distance="bray")
plot_ordination(physeq.day0, day0.ord, color="Type", shape="Soil") +
  theme_test()
```

**Statistics**

```{r}

```

### All data

I'd just like to get a feel for how the communities differed between treatments and over time.

Compare between treatments:

```{r}
physeq.ord = ordinate(physeq, method="NMDS", distance="bray")

# Compare soils
plot_ordination(physeq, physeq.ord, color="Soil") +
  facet_grid(~Amendment) +
  theme_test() +
  theme(strip.background = element_blank(),
        legend.title = element_blank())

plot_ordination(physeq, physeq.ord, color="Amendment") +
  facet_grid(~Soil) +
  theme_test() +
  theme(strip.background = element_blank(),
        legend.title = element_blank())
```

Over time:

```{r}
# Over time
plot_ordination(physeq, physeq.ord, color="Soil") +
  facet_grid(Amendment~Day) +
  theme_test() +
  theme(strip.background = element_blank(),
        legend.title = element_blank())
```

The communities do seem to separate over time, succesional moves further up, cropped moves further down. Moreso, in C-amended vs water treatments, maybe?
