---
title: "Dunlop - rarefy"
author: "Cassi Wattenburger"
date: "2/12/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(results = "show")
knitr::opts_chunk$set(message = FALSE)
```

# Import libraries

```{r}
library("phyloseq")
library("ape")
library("vegan")
library("ggplot2")
library("reshape")
library("dplyr")

sessionInfo()
```

# Import raw data

```{r}
# import count table, taxonomy table, tree, metadata
# growthrate.counts.finalv2.tsv is the same as the original, but I fixe a # in the header that screwed things up
otu <- read.table(file="/home/cassi/dunlop_popdynamics/output/growthrate.counts.finalv2.tsv", sep = '\t', header=TRUE)
tax <- read.table(file="/home/cassi/dunlop_popdynamics/output/growthrate.taxonomy.final.tsv", sep = '\t', header=TRUE)
meta <- read.table(file="../metadata/growthrate_metadata.tsv", sep = '\t', header=TRUE)
tree <- read_tree("/home/cassi/dunlop_popdynamics/output/growthrate.tree.final.nwk")
```

```{r}
# time point info for microcosm data
tp <- read.csv("../metadata/growthrate_timepoint_metadata.csv")
meta <- merge(meta, tp, by=1)

# experiment metadata (field sample or microcosm sample?)
meta <- mutate(meta, Experiment=(ifelse(is.na(Cammend), paste("Field"), paste("Microcosm"))))
```

```{r}
# remove sequencing controls
otu.noctrl <- otu[, -which(names(otu) %in% c("X261","X262","X263","X264","X265","X266","X267","X268","X269","X270",
                                             "X271","X272","X273","X274","X275","X276","X277","X278","X279","X280",
                                             "X281","X282","X283","X284","X285","X286","X287"))]
```

# Compare min/max read depth

```{r, , results="hide"}
# see the ratio of largest to smallest seq depth
# min and max sequencing depth
rownames(otu.noctrl) <- otu.noctrl[,1]
otu.noctrl <- otu.noctrl[,-1]
otu.noctrl.trans <- t(otu.noctrl)
depth <- rowSums(otu.noctrl.trans)
min(depth)
max(depth)
min(depth)/max(depth)
```

That's a huge desparity in read depth. Rarefication may be the best solution compared to CSS normalization; CSS normalization can't correct for the correlation b/w richness and read depth, but rarefication can.

```{r}
# calculate richness of each sample
otu.noctrl.trans <- t(otu.noctrl)
rich <- specnumber(otu.noctrl.trans) 
depth.df <- as.data.frame(depth)
rich.df <- as.data.frame(rich)
depth.rich <- merge(depth.df, rich.df, by=0)
```

# Richness by read depth bias
```{r}
ggplot(depth.rich, aes(x=depth, y=rich)) +
  geom_point() +
  labs(title="Raw count data", x="Read depth", y="Richness") +
  geom_text(aes(label=Row.names), hjust=0, vjust=0, size=2.5, color="red") +
  theme_test()
```

I'll rarefy the data.

# Create phyloseq object
```{r, , results="hide"}
# Reformat otu count, taxonomy, metadata dataframes so that phyloseq accepts them
rownames(otu) <- otu[,1]
otu <- otu[,-1]
#rownames(otu) <- sort(rownames(otu)) #!!!!!!!!!!!!! THIS IS A PROBLEM, doesn't sort rest of dataframe!!!
colnames(otu) <- gsub("X", "sa", colnames(otu))
otu.m <- as.matrix(otu)
rownames(tax) <- tax[,1]
tax <- tax[,-1] 
#rownames(tax) <- sort(rownames(tax)) #!!!!!!!!!!!!! THIS IS A PROBLEM, doesn't sort rest of dataframe!!!
tax.m <- as.matrix(tax)
rownames(meta) <- meta[,1]

# Create phyloseq object
OTU <- otu_table(otu.m,taxa_are_rows=TRUE)
TAX <- tax_table(tax.m)
SAM <- sample_data(meta, errorIfNULL=TRUE)
TREE <- tree
physeq.raw <- phyloseq(OTU, TAX, TREE, SAM)
```

Raw data phyloseq object:

```{r}
physeq.raw
```

```{r, eval=FALSE}
save(physeq.raw, file="rdata.files/gr_physeq.raw.RData")
```

# Remove non-prokaryotic ASVs

Remove controls first:

```{r}
# Remove controls from phyloseq object
all.raw = sample_names(physeq.raw)
remove.ctrls = c("sa261","sa262","sa263","sa264","sa265","sa266","sa267","sa268","sa269","sa270", "sa271","sa272","sa273","sa274","sa275","sa276","sa277","sa278","sa279","sa280","sa281","sa282","sa283","sa284","sa285","sa286","sa287")
keep.raw = all.raw[!(all.raw %in% remove.ctrls)]
physeq.raw.noctrl = subset_samples(physeq.raw, rownames(sample_data(physeq.raw)) %in% keep.raw)
```

Remove chloroplast, mitochondrial, unknown, and Eukaryotic domain sequences:

```{r, , results="hide"}
# Remove chloroplasts, mitochondria, unknown domains
chloro = subset_taxa(physeq.raw.noctrl, Class=="Chloroplast")
mito = subset_taxa(physeq.raw.noctrl, Family=="Mitochondria")
unknown = subset_taxa(physeq.raw.noctrl, Domain=="Unassigned")
unknown2 = subset_taxa(physeq.raw.noctrl, Domain=="putative Unassigned")
unknown3 = subset_taxa(physeq.raw.noctrl, Domain=="putative Bacteria")
eukarya = subset_taxa(physeq.raw.noctrl, Domain=="Eukaryota")
eukarya2 = subset_taxa(physeq.raw.noctrl, Domain=="putative Eukaryota")

# Remove non-prokarotes from dataset
allnames = taxa_names(physeq.raw.noctrl)
chloronames = taxa_names(chloro)
mitonames = taxa_names(mito)
unknownnames = taxa_names(unknown)
unknownnames2 = taxa_names(unknown2)
unknownnames3 = taxa_names(unknown3)
euknames = taxa_names(eukarya)
euknames2 = taxa_names(eukarya2)
keeptaxa1 = allnames[!(allnames %in% chloronames)]
keeptaxa2 = keeptaxa1[!(keeptaxa1 %in% mitonames)]
keeptaxa3 = keeptaxa2[!(keeptaxa2 %in% unknownnames)]
keeptaxa4 = keeptaxa3[!(keeptaxa3 %in% unknownnames2)]
keeptaxa5 = keeptaxa4[!(keeptaxa4 %in% unknownnames3)]
keeptaxa6 = keeptaxa5[!(keeptaxa5 %in% euknames)]
keeptaxa7 = keeptaxa6[!(keeptaxa6 %in% euknames2)]
# keeptaxa7 has all the taxa names that we will keep for analysis

physeq.raw.pruned = prune_taxa(keeptaxa7, physeq.raw.noctrl)
physeq.raw.pruned
```

```{r, eval=FALSE}
save(physeq.raw.pruned, file="rdata.files/gr_physeq.raw.pruned.RData")
```

# Rarefy

```{r, , results="hide"}
load("rdata.files/gr_physeq.raw.pruned.RData")

# look at read depth distribution by sample
reads.raw = data.frame(nreads = sample_sums(physeq.raw.pruned))
sam = data.frame(sample_data(physeq.raw.pruned))
read.sam <- merge(reads.raw, sam, by=0)
```

```{r}
ggplot(read.sam, aes(x=reorder(SampleID, -nreads), y=nreads)) +
  geom_bar(stat="identity") +
  ggtitle("Read depth by sample") +
  labs(x="Sample") +
  geom_hline(yintercept=10000) +
  theme_test()
```

Black line is at 10K reads.

Samples under 5k-12k reads:

```{r}
nrow(subset(read.sam, nreads<5000))
nrow(subset(read.sam, nreads<6000))
nrow(subset(read.sam, nreads<7000))
nrow(subset(read.sam, nreads<8000))
nrow(subset(read.sam, nreads<9000))
nrow(subset(read.sam, nreads<10000))
nrow(subset(read.sam, nreads<11000))
nrow(subset(read.sam, nreads<12000))
```

I'll rarefy to 9,000 taxa per sample, and lose 1 sample.

Rarefied dataset:

```{r}
physeq.rarefy = rarefy_even_depth(physeq.raw.pruned, 9000, rngseed=2018)
```

Sample 16 was lost due to low read depth.

```{r, eval=FALSE}
save(physeq.rarefy, file="rdata.files/gr_physeq.rarefy.RData")
```
