---
title: "Dunlop growth rate - descriptive statistics"
author: "Cassandra Wattenburger"
date: "2/13/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(results = "show")
```

**Goal of this script:** Explore the properties of the data and descriptive statistics of in-situ growth rates.

# Import libraries

```{r}
library(tidyverse)
library(cowplot)
library(phyloseq)
library(lmerTest)
library(emmeans)

sessionInfo()

rm(list=ls())
```

# Whole communities

**Number of taxa in whole dataset**

```{r}
# Import phyloseq object
load("../rdata.files/gr_physeq.rarefy.RData")

# Extract from phyloseq
count <-  data.frame(otu_table(physeq.rarefy))
tax <- data.frame(tax_table(physeq.rarefy)) %>%
  rownames_to_column(var="ASV")
meta <- data.frame(sam_data(physeq.rarefy))

# Remove unused samples
## Only microcosm samples relevant
meta.rm <- meta %>% filter(Experiment == "Microcosm")

# Merge count and metadata
count.meta <- count %>%
  t() %>% # flip data frame
  as.data.frame() %>%
  merge(meta.rm, by=0) %>%
  column_to_rownames(var="Row.names") %>%
  select(everything(), -(SampleID:Experiment))

# Remove absent ASVs
count.meta.rm0 <- colSums(count.meta, dims=1)
count.meta.rm0 <- count.meta.rm0[count.meta.rm0 > 0]

# Total ASVs
whole.asvs  <- names(count.meta.rm0)

# Total phyla
tax.rm0 <- tax[tax$ASV %in% whole.asvs,]
whole.phyla <- unique(tax.rm0$Phylum)
```

Number of ASVs:

```{r}
length(whole.asvs)
```

Total phyla:

```{r}
length(whole.phyla)
```

# Normalized microcosm and field data

```{r}
# Import normalized microcosm data
ucosm <- readRDS("../rdata.files/gr_ucosm.norm.clean.rds")
```

```{r, eval=False}
# Import normalized field data
field <- readRDS("../rdata.files/gr_field.norm.clean.rds") %>% 
  filter(norm_abund != "AsIs") %>% 
  mutate(norm_abund = as.numeric(norm_abund))

# total ASVs (not including 0 abundance)
ucosm %>% bind_rows(field) %>% 
  filter(norm_abund != 0) %>% # remove absent taxa
  distinct(ASV) %>% 
  nrow() 
```

Agrees with above.

**Most abundant phyla**

In micrococms at tp0:

```{r}
# Microcosms (tp0)
ucosm %>% 
  filter(Day==0) %>%
  group_by(Replicate, Phylum) %>% 
  summarize(total_abund = sum(norm_abund)) %>%
  ggplot(aes(y=fct_reorder(Phylum, -total_abund), x=total_abund)) +
  geom_boxplot() +
  labs(title="Microcosms at TP0", x="Total normalized abundance", y="Phylum") +
  theme_test()
```

In field:

```{r, eval=FALSE}
# Field
field %>% 
  group_by(Replicate, Phylum) %>% 
  summarize(total_abund = sum(norm_abund)) %>%
  ggplot(aes(y=fct_reorder(Phylum, -total_abund), x=total_abund)) +
  geom_boxplot() +
  labs(title="Field", x="Total normalized abundance", y="Phylum") +
  theme_test()
```

Not sure why that didn't reorder correctly.

# Growth

```{r}
# Import data
growth <- readRDS("../rdata.files/gr_gr.paprica.clean.rds")
```

Total ASVs with estimates:

```{r}
# Estimated growth rates
growth %>% distinct(ASV) %>% nrow() # total ASVs with estimates
```

Total phyla with estimated representative(s):

* Putative/unclassified taxa rolled into nearest related phylum

```{r}
growth %>% # total phyla, not counting 0 abundance ASVs, not counting putative/unclassified
  mutate(Phylum = str_remove(Phylum, "putative "), Phylum = str_remove(Phylum, "unclassified ")) %>% 
  distinct(Phylum) %>% 
  nrow()
```

Number of estimates per phylum:

```{r}
# Number of estimates per phylum
growth %>% 
  mutate(Phylum = str_remove(Class, "putative ")) %>% # roll putative into the phylum
  distinct(ASV, Phylum) %>%
  group_by(Phylum) %>%
  summarize(total = n()) %>%
  arrange(-total)

# Proteobacteria by class
growth %>% 
  filter(Phylum=="Proteobacteria") %>% 
  mutate(Class = str_remove(Class, "putative ")) %>% # roll putative into the class
  distinct(ASV, Class) %>%
  group_by(Class) %>%
  summarize(total = n()) %>%
  arrange(-total)
```

Visualize by replicate:

```{r}
growth %>% 
  group_by(Replicate, Phylum) %>%
  summarize(num_ASV = length(ASV)) %>%
  ggplot(aes(y=fct_reorder(Phylum, -num_ASV), x=num_ASV)) +
  geom_boxplot() +
  labs(title="Growth estimates", x="Total ASVs", y="Phylum") +
  theme_test()
```

Number of estimates per treatment:

```{r}
# Number of estimates per treatment
growth %>%
  group_by(Soil, Amendment) %>%
  distinct(ASV) %>%
  summarize(total = n())
```

# Death

```{r}
# Import data
death <- readRDS("../rdata.files/gr_death_estimates.rds")
```


```{r}
# Number of estimates per phylum
death %>% 
  mutate(Phylum = str_remove(Phylum, "putative ")) %>% # roll putative into the phylum
  distinct(ASV, Phylum) %>%
  group_by(Phylum) %>%
  summarize(total = n()) %>%
  arrange(-total)

# Proteobacteria by class
death %>% 
  filter(Phylum=="Proteobacteria" | Phylum=="unclassified Proteobacteria") %>% 
  mutate(Class = str_remove(Class, "putative ")) %>% # roll putative into the phylum
  distinct(ASV, Class) %>%
  group_by(Class) %>%
  summarize(total = n()) %>%
  arrange(-total)
```





