---
title: "Dunlop growth rates - groups"
author: "Cassandra Wattenburger"
date: "3/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(results = "show")
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

**Goal of this script:** Growth rate distributions seem bimodal. See if groups can be explicitly modeled and explore relationships with other variables.

Exploratory analysis.


# Import libraries

```{r}
library(tidyverse)
library(mclust)

sessionInfo()

rm(list=ls())

set.seed(2) # mclust uses EM algorhythm which is iterative and random
```


# Import data

```{r}
growth <- readRDS("../rdata.files/gr_gr.paprica.clean.rds")
```

Average across replicates to get a single estimated growth rate for each taxa in each treatment.

```{r}
# Average across replicates for each ASV
growth.asv <- growth %>%
  group_by(Soil, Amendment, ASV) %>%
  summarize(k = mean(k),
            start_day = mean(start_day),
            end_day = mean(end_day),
            start_abund = mean(start_abund),
            end_abund = mean(end_abund))

# Including PAPRICA results
# Have to remove archaea because they weren't predicted by program (NAs)
growth.paprica.asv <- growth %>%
  na.omit() %>%
  group_by(Soil, Amendment, ASV) %>%
  summarize(k = mean(k),
            start_day = mean(start_day),
            end_day = mean(end_day),
            start_abund = mean(start_abund),
            end_abund = mean(end_abund),
            n16S = mean(n16S),
            genome_size = mean(genome_size))
```


Visualize distributions:

```{r}
# Averaged across replicates
growth.asv %>%
  ggplot(aes(x=log(k), color=Soil)) +
  geom_density() +
  facet_wrap(~Amendment) +
  theme_test()
```


# Fit models and cluster


### Successional water control

```{r}
# Subset
S17n <- growth.asv %>% 
  filter(Soil=="S17" & Amendment=="N")

# Fit univariate gaussian mixture model
S17n.mclust <- densityMclust(log(S17n$k))
summary(S17n.mclust)
plot(S17n.mclust$BIC)
S17n.mclust$parameters
plot(S17n.mclust, what = "density", data = log(S17n$k))

# Bootstrap CIs of parameters
S17n.boot <- MclustBootstrap(S17n.mclust, nboot=999, type="bs", verbose=FALSE)
summary(S17n.boot, what = "ci")
plot(S17n.boot, what = "mean")

# Add classifications
S17n.class <- S17n %>%
  add_column(classification = S17n.mclust$classification) %>%
  mutate(classification = as_factor(classification))

S17n.class %>% 
  ggplot(aes(x=log(k), color=classification)) +
  geom_histogram() +
  theme_test()
```


### Cropped water control

```{r}
# Subset
C3n <- growth.asv %>% 
  filter(Soil=="C3" & Amendment=="N")

# Fit univariate gaussian mixture model
C3n.mclust <- densityMclust(log(C3n$k))
summary(C3n.mclust)
plot(C3n.mclust$BIC)
C3n.mclust$parameters
plot(C3n.mclust, what = "density", data = log(C3n$k))

# Bootstrap CIs of parameters
C3n.boot <- MclustBootstrap(C3n.mclust, nboot=999, type="bs", verbose=FALSE)
summary(C3n.boot, what = "ci")
plot(C3n.boot, what = "mean")

# Add classifications
C3n.class <- C3n %>%
  add_column(classification = C3n.mclust$classification) %>%
  mutate(classification = as_factor(classification))

C3n.class %>% 
  ggplot(aes(x=log(k), color=classification)) +
  geom_histogram() +
  theme_test()
```


### Successional C-amended

```{r}
# Subset
S17y <- growth.asv %>% 
  filter(Soil=="S17" & Amendment=="Y")

# Fit univariate gaussian mixture model
S17y.mclust <- densityMclust(log(S17y$k))
summary(S17y.mclust)
plot(S17y.mclust$BIC)
S17y.mclust$parameters
plot(S17y.mclust, what = "density", data = log(S17y$k))

# Bootstrap CIs of parameters
S17y.boot <- MclustBootstrap(S17y.mclust, nboot=999, type="bs", verbose=FALSE)
summary(S17y.boot, what = "ci")
plot(S17y.boot, what = "mean")

# Add classifications
S17y.class <- S17y %>%
  add_column(classification = S17y.mclust$classification) %>%
  mutate(classification = as_factor(classification))

S17y.class %>% 
  ggplot(aes(x=log(k), color=classification)) +
  geom_histogram() +
  theme_test()
```


### Cropped C-amended

```{r}
# Subset
C3y <- growth.asv %>% 
  filter(Soil=="C3" & Amendment=="Y")

# Fit univariate gaussian mixture model
C3y.mclust <- densityMclust(log(C3y$k))
summary(C3y.mclust)
plot(C3y.mclust$BIC)
C3y.mclust$parameters
plot(C3y.mclust, what = "density", data = log(C3y$k))

# Add classifications
C3y.class <- C3y %>%
  add_column(classification = C3y.mclust$classification) %>%
  mutate(classification = as_factor(classification))
```

It looks like a model with a single distribution best fits this treatment, rather than 2 distributions.

# Figure

Code to create and plot clusters borrowed from https://github.com/jlw-ecoevo/eggo/blob/master/Scripts/RefSeq_Analysis.R

```{r}
# Generate clusters for graphing, scaled to proportion
C3n.clusters <- data.frame(x = seq(-5,0.5,0.01),
                    cl1 = dnorm(seq(-5,0.5,0.01),
                              mean = C3n.mclust$parameters$mean[1],
                              sd = sqrt(C3n.mclust$p$variance$sigmasq))*C3n.mclust$parameters$pro[1],
                    cl2 = dnorm(seq(-5,0.5,0.01),
                              mean = C3n.mclust$parameters$mean[2],
                              sd = sqrt(C3n.mclust$p$variance$sigmasq))*C3n.mclust$parameters$pro[2])

# Plot
plot1 <- growth.asv %>%
  filter(Soil=="C3", Amendment=="N") %>%
  ggplot() +
  geom_polygon(data=C3n.clusters, aes(x=x, y=cl1), linetype=2, alpha=0.6, color="black") +
  geom_polygon(data=C3n.clusters, aes(x=x, y=cl2), linetype=2, alpha=0.6, color="black") +
  geom_density(aes(x=log(k))) +
  labs(title="Cropped, water", y="Kernel density", x="Specific growth rate (units), ln") +
  theme_classic() +
  theme(plot.title = element_text(size=9),
        text = element_text(size=9),
        strip.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.title = element_blank())
plot1
```

```{r}
# Generate clusters for graphing, scaled to proportion
C3y.clusters <- data.frame(x = seq(-5,0.5,0.01),
                    cl1 = dnorm(seq(-5,0.5,0.01),
                              mean = C3y.mclust$parameters$mean[1],
                              sd = sqrt(C3y.mclust$parameters$variance$sigmasq)))

# Plot
plot2 <- growth.asv %>%
  filter(Soil=="C3", Amendment=="Y") %>%
  ggplot() +
  geom_polygon(data=C3y.clusters, aes(x=x, y=cl1), linetype=2, alpha=0.6, color="black") +
  geom_density(aes(x=log(k))) +
  labs(title="Cropped, resources", y="Kernel density", x="Specific growth rate (units), ln") +
  theme_classic() +
  theme(plot.title = element_text(size=9),
        text = element_text(size=9),
        strip.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.title = element_blank())
plot2
```

```{r}
# Generate clusters for graphing, scaled to proportion
S17n.clusters <- data.frame(x = seq(-5,0.5,0.01),
                    cl1 = dnorm(seq(-5,0.5,0.01),
                              mean = S17n.mclust$parameters$mean[1],
                              sd = sqrt(S17n.mclust$p$variance$sigmasq[1]))*S17n.mclust$parameters$pro[1],
                    cl2 = dnorm(seq(-5,0.5,0.01),
                              mean = S17n.mclust$parameters$mean[2],
                              sd = sqrt(S17n.mclust$p$variance$sigmasq[2]))*S17n.mclust$parameters$pro[2])

# Plot
plot3 <- growth.asv %>%
  filter(Soil=="S17", Amendment=="N") %>%
  ggplot() +
  geom_polygon(data=S17n.clusters, aes(x=x, y=cl1), linetype=2, alpha=0.6, color="black") +
  geom_polygon(data=S17n.clusters, aes(x=x, y=cl2), linetype=2, alpha=0.6, color="black") +
  geom_density(aes(x=log(k))) +
  labs(title="Successional, water", y="Kernel density", x="Specific growth rate (units), ln") +
  theme_classic() +
  theme(plot.title = element_text(size=9),
        text = element_text(size=9),
        strip.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.title = element_blank())
plot3
```

```{r}
# Generate clusters for graphing, scaled to proportion
S17y.clusters <- data.frame(x = seq(-5,0.5,0.01),
                    cl1 = dnorm(seq(-5,0.5,0.01),
                              mean = S17y.mclust$parameters$mean[1],
                              sd = sqrt(S17y.mclust$p$variance$sigmasq[1]))*S17y.mclust$parameters$pro[1],
                    cl2 = dnorm(seq(-5,0.5,0.01),
                              mean = S17y.mclust$parameters$mean[2],
                              sd = sqrt(S17y.mclust$p$variance$sigmasq[2]))*S17y.mclust$parameters$pro[2])

# Plot
plot4 <- growth.asv %>%
  filter(Soil=="S17", Amendment=="Y") %>%
  ggplot() +
  geom_polygon(data=S17y.clusters, aes(x=x, y=cl1), linetype=2, alpha=0.6, color="black") +
  geom_polygon(data=S17y.clusters, aes(x=x, y=cl2), linetype=2, alpha=0.6, color="black") +
  geom_density(aes(x=log(k))) +
  labs(title="Successional, resources", y="Kernel density", x="Specific growth rate (units), ln") +
  theme_classic() +
  theme(plot.title = element_text(size=9),
        text = element_text(size=9),
        strip.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.title = element_blank())
plot4
```

```{r}
figure <- plot_grid(plot1, plot2, plot3, plot4)
figure

ggsave(figure, filename="../figures/fig_groups.png", units="mm", width=85, height=85)
```

```{r}
plot(C3n.mclust$BIC) +
title("Cropped, water")
plot(S17n.mclust$BIC) +
title("Succesional, water")
plot(C3y.mclust$BIC) +
title("Cropped, resources")
plot(S17y.mclust$BIC) +
title("Succesional, water")

```
