---
title: "Untitled"
author: "Cassandra Wattenburger"
date: "12/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(results = "show")
knitr::opts_chunk$set(message = FALSE)
```

```{r}
rm(list=ls())

library("tidyverse")

sessionInfo()
```

# Import data

```{r}
# Growth and death estimates
growth <- readRDS("../rdata.files/gr_gr.paprica.clean.rds") %>% 
  rename(growth_start_day=start_day, growth_end_day=end_day,
         growth_start_abund=start_abund, growth_end_abund=end_abund) %>% 
    mutate(Label = paste0(Soil, Amendment, Replicate, ASV),
         growth_change_abund = growth_end_abund - growth_start_abund,
         growth_length = growth_end_day - growth_start_day)

death <- readRDS("../rdata.files/gr_death_estimates_after.rds") %>% 

  select(everything(), -start_pt, -end_pt, -change_abund) %>% 
  rename(death_start_day=start_day, death_end_day=end_day,
         death_start_abund=start_abund, death_end_abund=end_abund) %>% 
  mutate(death_change_abund = death_end_abund - death_start_abund,
         death_length = death_end_day - death_start_day)

# Normalized abundances
norm <- readRDS("../rdata.files/gr_norm_growth_estimated.rds") %>% 
  mutate(abund_ln = log(norm_abund))

# Taxa with estimated growth and then death
growth_death <- inner_join(growth, death) %>% 
  select(Label, Soil, Amendment, Replicate, Domain:Genus, ASV,
         k, g, growth_start_day, growth_end_day, growth_start_abund, growth_end_abund, growth_change_abund, growth_length,
         death_rate, death_start_day, death_end_day, death_start_abund, death_end_abund, death_change_abund, death_length,
         edge_num:n16S)
```

# Graph growth and death estimates on timeseries

```{r}
# For messing with graphs
testing <- head(growth_death)

# Taxonomy
tax <- select(growth_death, ASV, Domain:Genus) %>% unique()

# Plot all
for (l in as.character(growth_death$Label)) {
  # Subset time series
  est_label <- filter(growth_death, Label==l)
  norm_label <- filter(norm, Label==l) %>% 
    arrange(Day)
  # Title information
  asv <- est_label$ASV
  tax_info <- filter(tax, ASV == asv)
  title <- paste0(as.character(tax_info$Phylum), ", ", as.character(tax_info$Genus))
  # Graph with estimate
  graph <- ggplot(norm_label, aes(x=Day, y=abund_ln)) +
    geom_point(shape=1, size=2, color="#6F7378") +
    geom_line(color="#6F7378") +
    # Growth estimate
    geom_smooth(method="lm", data=filter(norm_label, Day >= est_label$growth_start_day & Day <= est_label$growth_end_day),
                linetype=2, color="green") +
    # Death estimate
    geom_smooth(method="lm", data=filter(norm_label, Day >= est_label$death_start_day & Day <= est_label$death_end_day),
                linetype=2, color="red") +
    # Formatting
    labs(title=title, x="Day", y="ln ARNIS ratio") +
    theme_test() +
    theme(title = element_text(size=18),
        axis.title = element_text(size=16),
        axis.text = element_text(size=14))
  print(graph)
}

```

Due to sparser sampling towards end of incubations, death estimates are much less certain/patchy compared to growth estimates. Less data was used for death estimates as well, which means the false positive rate was lower and the p-value threshold higher.

# Treatment-level

Average across ASVs to get a treatment-level estimate for parameters in each replicate. Allows a broad understanding of how soil treatment affected the growing/dying community as a whole.

```{r}
# Summarize parameters for each replicate in each treatment
growth_death_trt <- growth_death %>% 
  group_by(Soil, Amendment, Replicate) %>% 
    summarize(k = mean(k), g = mean(g),
            growth_start_day = mean(growth_start_day), growth_end_day = mean(growth_end_day), 
            growth_length = mean(growth_length),
            growth_start_abund = mean(growth_start_abund), growth_end_abund = mean(growth_end_abund), 
            growth_change_abund = mean(growth_change_abund),
            growth_start_abund_corrected = mean(growth_start_abund/n16S, na.rm = TRUE), # 16S copy num correction
            growth_change_abund_corrected = mean(growth_change_abund/n16S, na.rm = TRUE), # 16S copy num correction
            death_rate = mean(death_rate),
            death_start_day = mean(death_start_day), death_end_day = mean(death_end_day), 
            death_length = mean(death_length),
            death_start_abund = mean(death_start_abund), death_end_abund = mean(death_end_abund), 
            death_change_abund = mean(death_change_abund),
            death_start_abund_corrected = mean(death_start_abund/n16S, na.rm = TRUE), # 16S copy num correction
            death_change_abund_corrected = mean(death_change_abund/n16S, na.rm = TRUE)) %>% # 16S copy num correction
  ungroup()
```

### Number of taxa that grew then died vs just grew

```{r}
# Tally
count_growth_death <- growth_death %>% 
  group_by(Soil, Amendment, Replicate) %>% 
  summarize(count = n()) %>% 
  ungroup()

count_growth_death

# Graph
count_growth_death %>% 
  ggplot(aes(x=Soil, y=count, color=Amendment)) +
  geom_point() +
  theme_test()
```

General trend of more estiamted growth and death in amended compared to unamended.

**Statistics**

Linear model

```{r}
growth_death_count_lm <- lm(log(count) ~ Soil*Amendment, data=count_growth_death)
hist(resid(growth_death_count_lm))
plot(predict(growth_death_count_lm), resid(growth_death_count_lm))
```

Quite unequal variances, will use Welch's t-test to test amendment effect.

Welch's t-test

* Untransformed data

```{r}
growth_death_count_welch <- t.test(count ~ Amendment, data=count_growth_death, var.equal=FALSE) 
growth_death_count_welch
```

### Death rate

```{r}
growth_death_trt %>% 
  ggplot(aes(x=Soil, y=death_rate, color=Amendment)) +
  geom_point() +
  theme_test()
```

### Death start day

```{r}
growth_death_trt %>% 
  ggplot(aes(x=Soil, y=death_start_day, color=Amendment)) +
  geom_point() +
  theme_test()
```

### Death length

```{r}
growth_death_trt %>% 
  ggplot(aes(x=Soil, y=death_length, color=Amendment)) +
  geom_point() +
  theme_test()
```

### Death change abundance

```{r}
growth_death_trt %>% 
  ggplot(aes(x=Soil, y=death_change_abund, color=Amendment)) +
  geom_point() +
  theme_test()
```

Seems like trend of greater changes in abundance due to death in amended soils?

**Statistics**

Linear model

```{r}
death_chabund_trt_lm <- lm(log(abs(death_change_abund)) ~ Soil+Amendment, data=growth_death_trt)
hist(resid(death_chabund_trt_lm))
plot(predict(death_chabund_trt_lm), resid(death_chabund_trt_lm))
```

I don't like the unequal variance here, even with a log transform. I'll use a Welch's t-test to test Amendment effect instead.

Welch's t-test

* Non-transformed data

```{r}
death_chabund_trt_welch <- t.test(death_change_abund ~ Amendment, data=growth_death_trt, var.equal=FALSE)
death_chabund_trt_welch
```

16S copy number corrected:

```{r}
growth_death_trt %>% 
  ggplot(aes(x=Soil, y=death_change_abund_corrected, color=Amendment)) +
  geom_point() +
  theme_test()
```

**Statistics**

Linear model

```{r}
death_chabund_corr_trt_lm <- lm(log(abs(death_change_abund_corrected)) ~ Soil+Amendment, data=growth_death_trt)
hist(resid(death_chabund_corr_trt_lm))
plot(predict(death_chabund_corr_trt_lm), resid(death_chabund_corr_trt_lm))
```

I don't like the unequal variance here, even with a log transform. I'll use a Welch's t-test to test Amendment effect instead.

Welch's t-test

* Non-transformed data

```{r}
death_chabund_corr_trt_welch <- t.test(death_change_abund_corrected ~ Amendment, data=growth_death_trt, var.equal=FALSE)
death_chabund_corr_trt_welch
```

The trend holds.

# Population level

Average each ASV across replicates to get a parameter estimate for each taxa in each treatment. This allows a more nuanced investigation of how soil treatments interact with growth responses of individual ASVs/populations rather than the community as a whole.

```{r}
# Average growth estimates for each ASV
growth_asv <- growth %>% 
  group_by(ASV, Soil, Amendment) %>% 
  summarize(k = mean(k), g = mean(g),
            growth_start_day = mean(growth_start_day), growth_end_day = mean(growth_end_day), 
            growth_length = mean(growth_length),
            growth_start_abund = mean(growth_start_abund), growth_end_abund = mean(growth_end_abund), 
            growth_change_abund = mean(growth_change_abund),
            growth_start_abund_corrected = mean(growth_start_abund/n16S, na.rm=TRUE),
            growth_change_abund_corrected = mean(growth_change_abund/n16S, na.rm=TRUE),
            n16S = mean(n16S, na.rm=TRUE)) %>% 
  ungroup()

# Average death estimates for each ASV
death_asv <- death %>% 
  group_by(ASV, Soil, Amendment) %>% 
  summarize(death_rate = mean(death_rate),
            death_start_day = mean(death_start_day), death_end_day = mean(death_end_day), 
            death_length = mean(death_length),
            death_start_abund = mean(death_start_abund), death_end_abund = mean(death_end_abund), 
            death_change_abund = mean(death_change_abund)) %>% 
  ungroup()

# Combined
growth_death_asv <- growth_death %>% 
  group_by(ASV, Soil, Amendment) %>% 
  summarize(k = mean(k), g = mean(g),
            growth_start_day = mean(growth_start_day), growth_end_day = mean(growth_end_day), 
            growth_length = mean(growth_length),
            growth_start_abund = mean(growth_start_abund), growth_end_abund = mean(growth_end_abund), 
            growth_change_abund = mean(growth_change_abund),
            growth_start_abund_corrected = mean(growth_start_abund/n16S, na.rm=TRUE),
            growth_change_abund_corrected = mean(growth_change_abund/n16S, na.rm=TRUE),
            death_rate = mean(death_rate),
            death_start_day = mean(death_start_day), death_end_day = mean(death_end_day), 
            death_length = mean(death_length),
            death_start_abund = mean(death_start_abund), death_end_abund = mean(death_end_abund), 
            death_change_abund = mean(death_change_abund),
            death_start_abund_corrected = mean(death_start_abund/n16S, na.rm=TRUE),
            death_change_abund_corrected = mean(death_change_abund/n16S, na.rm=TRUE),
            n16S = mean(n16S, na.rm=TRUE)) %>% 
  ungroup()
```

```{r}
# Presence/absence of death in growing ASVs
only_growth_asv <- anti_join(growth_asv, death_asv, by=c("ASV", "Soil", "Amendment")) %>% 
  add_column(died ="no")

growth_death_asv <- growth_death_asv %>% 
  add_column(died="yes")

# Include ASVs without detected death after growth
all_asv <- bind_rows(only_growth_asv, growth_death_asv) %>% 
  mutate()
```

### Growth rate vs death yes/no

Hypothesis: Taxa with faster growth rates more likely to experience death afterwards
* boom/bust lifestyle
* didn't sample enough/long enough to capture slower grower death

```{r}
# Graph
all_asv %>% 
  ggplot(aes(x=died, y=log(k))) +
  geom_boxplot() +
  geom_jitter(alpha=0.5) +
  facet_wrap(Soil~Amendment) +
  theme_test()
```

Greater variability and sample size of taxa that did not have stimated death after growth. Dificult to say whether this difference in variation is real or due to undersampling.

### Growth lag time vs death yes/no

Hypothesis: Taxa with earlier start day more likely to experience death afterwards
* boom/bust lifestyle
* didn't sample enough/long enough to capture death of later growing organisms

```{r}
# Graph
all_asv %>% 
  mutate(growth_start_day = growth_start_day + 0.1) %>% # Can't log transform 0, add a psuedo-value
  ggplot(aes(x=died, y=log(growth_start_day))) +
  geom_boxplot() +
  geom_jitter(alpha=0.5) +
  facet_wrap(Soil~Amendment) +
  theme_test()
```

See above interpretation.

### Growth change abundance vs death yes/no

Hypothesis: Taxa that grew more biomass more likely to experience death afterwards
* boom/bust lifestyle
* didn't sample enough/long enough to slower/less growth and death

```{r}
# Graph
all_asv %>% 
  ggplot(aes(x=died, y=log(growth_change_abund))) +
  geom_boxplot() +
  geom_jitter(alpha=0.5) +
  facet_wrap(Soil~Amendment) +
  theme_test()
```

Same interpretation. If there are any trends, they are slight.

### Growth rate vs death rate

Hypothesis: Taxa with faster growth rate more likely to experience faster death
* boom/bust lifestyle

```{r}
# Graphize
growth_death_asv %>% 
  ggplot(aes(x=k, y=log(abs(death_rate)))) +
  geom_smooth(method="lm", linetype=2, alpha=0.5) +
  geom_point(alpha=0.5) +
  facet_wrap(Soil~Amendment) +
  theme_test()
```

Perhaps a trend of slight positive relationship but too slight/variable to confirm.

### Growth lag time vs death rate

Hypothesis: Taxa with earlier start day more likely to experience faster death
* boom/bust lifestyle

```{r}
# Graph
growth_death_asv %>% 
  ggplot(aes(x=growth_start_day, y=death_rate)) +
  geom_point(alpha=0.5) +
  geom_smooth(method="lm", linetype=2, alpha=0.5) +
  facet_wrap(Soil~Amendment) +
  theme_test()
```

Not much trend.

# Growth change in abundance vs death rate

Hypothesis: Taxa that grew more more likely to experience faster death
* boom/bust lifestyle

```{r}
# Graphize
growth_death_asv %>% 
  ggplot(aes(x=log(growth_change_abund), y=log(abs(death_rate)))) + # cannot take ln of negative number
  geom_point(alpha=0.5) +
  geom_smooth(method="lm", linetype=2, alpha=0.5) +
  facet_wrap(Soil~Amendment) +
  theme_test()
```

Possibly slight positive correlation in amended soils but negative in unamended? This would mean that in unamended soils, as the change in growth biomass increased, the death rate tended to be lower, but in the amended soils, the opposite was true.

**Statistics**

Pearson correlations with multiple test correction

```{r}
# Treatments
grchabund_deathrate_pcor_trts <- data.frame()
for (s in c("C3", "S17")) {
  for (a in c("Y", "N")) {
    data_sub <- filter(growth_death_asv, Soil==s & Amendment==a)
    pcor <- cor.test(log(data_sub$growth_change_abund), log(abs(data_sub$death_rate)), method="pearson")
    this_row <-  data.frame("Soil"=s, "Amendment"=a, "estimate"=pcor$estimate, "pvalue"=pcor$p.value)
    grchabund_deathrate_pcor_trts <- bind_rows(grchabund_deathrate_pcor_trts, this_row)
  }
}

# Multiple test correction
padj <- p.adjust(grchabund_deathrate_pcor_trts$pvalue, method="holm", n=4)
grchabund_deathrate_pcor_trts <- bind_cols(grchabund_deathrate_pcor_trts, padj=padj)

# Results
grchabund_deathrate_pcor_trts
```

No significance.

# Growth rate vs death change in abundance

Hypothesis: Taxa with faster growth rate lose more biomass to death
* boom/bust lifestyle

```{r}
# Graph
growth_death_asv %>% 
  ggplot(aes(x=k, y=log(abs(death_change_abund)))) +
  geom_point(alpha=0.5) +
  geom_smooth(method="lm", linetype=2, alpha=0.5) +
  facet_wrap(Soil~Amendment) +
  theme_test()
```

Positive correlation trend? This would mean that as growth rate increases, the amount of biomass lost to death increases.

**Statistics**

Pearson correlations with multiple test correction

```{r}
# Treatments
k_deathchabund_pcor_trts <- data.frame()
for (s in c("C3", "S17")) {
  for (a in c("Y", "N")) {
    data_sub <- filter(growth_death_asv, Soil==s & Amendment==a)
    pcor <- cor.test(log(data_sub$k), log(abs(data_sub$death_change_abund)), method="pearson")
    this_row <-  data.frame("Soil"=s, "Amendment"=a, "estimate"=pcor$estimate, "pvalue"=pcor$p.value)
    k_deathchabund_pcor_trts <- bind_rows(k_deathchabund_pcor_trts, this_row)
  }
}

# Multiple test correction
padj <- p.adjust(k_deathchabund_pcor_trts$pvalue, method="holm", n=4)
k_deathchabund_pcor_trts <- bind_cols(k_deathchabund_pcor_trts, padj=padj)

# Results
k_deathchabund_pcor_trts
```

Only significant relationship is in cropped, amended, with positive estimate. Similar trend in successional, amended.

### Growth start day vs death change in abundance

Hypothesis: Taxa that started earlier die off more
* boom/bust lifestyle

```{r}
# Graph
growth_death_asv %>% 
  ggplot(aes(x=growth_start_day, y=log(abs(death_change_abund)))) +
  geom_point(alpha=0.5) +
  geom_smooth(method="lm", linetype=2, alpha=0.5) +
  facet_wrap(Soil~Amendment) +
  theme_test()
```

No clear trends.

# Growth change in abundance vs death change in abundance

Hypothesis: Taxa that accumulated more biomass during growth can and will lose more biomass to death
* boom/bust lifestyle

```{r}
# Graph
growth_death_asv %>% 
  filter(died=="yes") %>%
  ggplot(aes(x=log(growth_change_abund), y=log(abs(death_change_abund)))) +
  geom_point(alpha=0.5) +
  geom_smooth(method="lm", linetype=2, alpha=0.5) +
  facet_wrap(Soil~Amendment) +
  theme_test()
```

Positive correlation between change in abundance due to death and growth? If a taxa grew more, it died more. There also seems to be a treatment effect but this could be due to undersampling things that grew then died in the unammended soil.

**Statistics**

Pearson correlations with multiple test correction

```{r}
# Treatments
chabund_growth_death_pcor <- data.frame()
for (s in c("C3", "S17")) {
  for (a in c("Y", "N")) {
    data_sub <- filter(growth_death_asv, Soil==s & Amendment==a)
    pcor <- cor.test(log(data_sub$growth_change_abund), log(abs(data_sub$death_change_abund)), method="pearson")
    this_row <-  data.frame("Soil"=s, "Amendment"=a, "estimate"=pcor$estimate, "pvalue"=pcor$p.value)
    chabund_growth_death_pcor <- bind_rows(chabund_growth_death_pcor, this_row)
  }
}

# Multiple test correction
padj <- p.adjust(chabund_growth_death_pcor$pvalue, method="holm", n=4)
chabund_growth_death_pcor_trts <- bind_cols(chabund_growth_death_pcor, padj=padj)

# Results
chabund_growth_death_pcor
```

Make sure 16S copy number does not affect results:

```{r}
# Graph
growth_death_asv %>% 
  ggplot(aes(x=log(growth_change_abund_corrected), y=log(abs(death_change_abund_corrected)))) +
  geom_point(alpha=0.5) +
  geom_smooth(method="lm", linetype=2, alpha=0.5) +
  facet_wrap(Soil~Amendment) +
  theme_test()
```

Trend actually seems clearer.

**Statistics**

Pearson correlations with multiple test correction

```{r}
# Treatments
chabund_growth_death_corr_pcor <- data.frame()
for (s in c("C3", "S17")) {
  for (a in c("Y", "N")) {
    data_sub <- filter(growth_death_asv, Soil==s & Amendment==a)
    pcor <- cor.test(log(data_sub$growth_change_abund_corrected), log(abs(data_sub$death_change_abund_corrected)), method="pearson")
    this_row <-  data.frame("Soil"=s, "Amendment"=a, "estimate"=pcor$estimate, "pvalue"=pcor$p.value)
    chabund_growth_death_corr_pcor <- bind_rows(chabund_growth_death_corr_pcor, this_row)
  }
}

# Multiple test correction
padj <- p.adjust(chabund_growth_death_corr_pcor$pvalue, method="holm", n=4)
chabund_growth_death_corr_pcor <- bind_cols(chabund_growth_death_corr_pcor, padj=padj)

# Results
chabund_growth_death_corr_pcor
```

Results the same.




